---
title: "Figure 5"
output: html_notebook
---

#1 Permutation 10000 times for each dyad

## 30-day loc vector for each id function
```{r}
vector_const_loc <- function(k){
  
  colony <- colonyname[k]
  id <- idlist[[k]]
  size <- length(id)
  dataset <- dayname[[k]]
  
  #prepare initial variable for each id
  for(i in 1:size){
    assign(paste0("d_",i), "0", envir = .GlobalEnv)
  }
  
  #30-day vector for each id
  for(name in dataset){
    d <- fread(paste0("data/processed_data/locnest_data/", colony, "/", name, ".csv")) %>% 
      mutate(Time1s = ceiling(Time-0.01)) %>% 
      distinct(Time1s, .keep_all = TRUE) %>% 
      select(-Time, -Time1s)
    for(i in 1:size){
      di <- d %>% pull(i) %>% replace_na("0")
      nth <- paste0("d_",i)
      assign(nth, c(get(nth), di), envir = .GlobalEnv)
    }
  }
  for(i in 1:size){
    nth <- paste0("d_",i)
    assign(nth, get(nth)[-1], envir = .GlobalEnv)
  }
}
```

##Permutation for location function
```{r}
##synchronized times calculation function (location)
syn_loc <- function(k, N, ran, ID1, ID2){
  if(ran[k] == 1){ID1r <- ID1}
  if(ran[k] != 1){ID1r <- ID1[c(ran[k]:N,1:(ran[k]-1))]}
  ID12 <- tibble(ID1r, ID2) %>% 
    mutate(Syn = case_when(ID1r == "0" & ID2 == "0" ~ 0, 
                           ID1r == ID2 ~ 1, 
                           TRUE ~ 0)) %>% 
    pull(Syn)
  Dl <<- sum(ID12 == 1)
}

#permutation (circular) test for rest/active function
syn_loc_pt.test <- function(ID1, ID2, nperm){
  #determine sequence to permute vector randomly
  N <- length(ID1)
  set.seed(20)
  ran <- sample(2:N, nperm, replace = FALSE)
  ran <- c(1, ran) #add actual value to the first element
  
  DL <- 0
  for(k in 1:(nperm+1)){
    syn_loc(k, N, ran, ID1, ID2)
    DL <- c(DL, Dl)
  }
  DL <<- DL[-1]
}

#permutation test for rest/active performing function
syn_loc_pttest <- function(colony, nperm){
  
  size <- idlist[colony] %>% unlist() %>% length()
  pb <- txtProgressBar(min = 1, max = size*(size-1)/2, style = 3)
  Dout <- 0
  DLs <- list(0)
  
  for(i in 1:(size-1)){
    for(j in (i+1):size){
      #progress bar
      n <- size*(size-1)/2 - (size-i+1)*(size-i)/2 + (j-i)
      setTxtProgressBar(pb, n)
      
      #30-day vectors for ID1 & ID2
      ID1 <- get(paste0("d_",i))
      ID2 <- get(paste0("d_",j))
      
      #extract both outside data
      d_ij <- tibble(ID1, ID2) %>% filter(ID1 != "N", ID2 != "N")
      ID1 <- d_ij$ID1
      ID2 <- d_ij$ID2
      #extract data size (ie outside times data)
      Dout <- c(Dout, nrow(d_ij))
      
      #permutation test
      syn_loc_pt.test(ID1, ID2, nperm)
      DLs[n] <- list(DL)
    }
  }
  D_out <<- Dout[-1]
  DL_s <<- DLs
}
```


##Run function
```{r}
nperm <- 9999

for(l in 1:5){
  vector_const_loc(l)
  system.time(syn_loc_pttest(l, nperm))
  
  size <- idlist[l] %>% unlist() %>% length()
  nth_out <- paste0("Dout_", l)
  assign(nth_out, D_out)
  DL_s <- data.frame(DL_s)
  colnames(DL_s) <- 1:(size*(size-1)/2)
  nth <- paste0("DL_", l)
  assign(nth, DL_s)
}
#output data: Dout_1~5, DL_1~5


#trial for the first dyad
#colony <- 1
#size <- idlist[colony] %>% unlist() %>% length()
#N <- length(d_1)
#ID1 <- get(paste0("d_",1))
#ID2 <- get(paste0("d_",2))
#d_ij <- tibble(ID1, ID2) %>% filter(ID1 != "N", ID2 != "N")
#Dout <- 0
#Dout <- c(Dout, nrow(d_ij))
#DLs <- list(0)
#set.seed(10)
#ran <- sample(2:N, nperm, replace = FALSE)
#ran <- c(1, ran)
#syn_loc_pt.test(ran, ID1, ID2, nperm)
#n <- size*(size-1)/2 - (size-1+1)*(size-1)/2 + (2-1)
#DLs[n] <- list(DL)
#DL_s <- DLs[[1]] %>% as.vector()
#DL_s <- tibble(DL_s)
```

##Integrate data
```{r}
trans_tidy_loc <- function(d, i){
  colony <- colonyname[i]
  d %>% 
    mutate(index = 1:10000) %>% 
    pivot_longer(cols = -index, names_to = "Comb", values_to = "L_val") %>% 
    mutate(Comb = as.integer(Comb)) %>% 
    arrange(Comb) %>% 
    mutate(Colony = colony) %>% 
    select(Colony, Comb, index, L_val)
}

DL_1_l <- trans_tidy_loc(DL_1, 1)
DL_2_l <- trans_tidy_loc(DL_2, 2)
DL_3_l <- trans_tidy_loc(DL_3, 3)
DL_4_l <- trans_tidy_loc(DL_4, 4)
DL_5_l <- trans_tidy_loc(DL_5, 5)

DL_t_l <- bind_rows(DL_1_l, DL_2_l, DL_3_l, DL_4_l, DL_5_l) %>% 
  left_join(d_out, join_by(Colony, Comb)) %>% 
  mutate(pL_val = L_val / Out)
```

##Dyad information
```{r}
ID_att <- read.csv("data/pre_info_data/individual_information.csv") %>% mutate(Birth = as.Date(Birth))

ID_att_1 <- ID_att %>% select(Colony, ID1 = ID, Caste1 = Caste, Sex1 = Sex, AgeD1 = AgeD, Weight1 = Weight)
ID_att_2 <- ID_att %>% select(Colony, ID2 = ID, Caste2 = Caste, Sex2 = Sex, AgeD2 = AgeD, Weight2 = Weight)

idattcomb_each <- function(colony, size){
  rep(ID_att %>% filter(Colony == colony) %>% pull(ID), each = size)
}
idattcomb <- function(colony, size){
  rep(ID_att %>% filter(Colony == colony) %>% pull(ID), size)
}

ID1 <- "0"; ID2 <- "0"; Colony <- "0"; ID1n <- 0; ID2n <- 0 
for(i in 1:5){
  colony <- colonyname[i]
  size <- idlist[i] %>% unlist() %>% length()
  
  ID1_i <- idattcomb_each(colony, size); ID2_i <- idattcomb(colony, size)
  Colony_i <- rep(colony, size^2)
  ID1n_i <- rep(1:size, each = size); ID2n_i <- rep(1:size, size)
  
  ID1 <- c(ID1, ID1_i); ID2 <- c(ID2, ID2_i)
  Colony <- c(Colony, Colony_i)
  ID1n <- c(ID1n, ID1n_i); ID2n <- c(ID2n, ID2n_i)
}
Comb <- c(1:231, 1:136, 1:171, 1:231, 1:231)

ID_att_comb <- tibble(Colony, ID1, ID2, ID1n, ID2n) %>% 
  slice(-1) %>% filter(ID1 != ID2) %>% 
  mutate(sum = ID1n + ID2n, dif = abs(ID1n - ID2n)) %>% 
  group_by(Colony) %>% distinct(sum, dif, .keep_all = T) %>% 
  ungroup() %>% mutate(Comb = Comb) %>% 
  select(Colony, Comb, ID1, ID2)
```


#2 Visualization of p-value & proximity index

##p-value data
```{r}
#p-value summarize function
pv_data <- function(d){
  d %>% group_by(Comb) %>% 
    summarise(Value = Syn[index == 1],
              Med = median(Syn),
              Pn = sum(Syn >= Syn[index == 1]) / 10000) %>% 
    mutate(Sign = if_else(Pn < 0.5, "Plus", "Minus")) %>% 
    mutate(P = if_else(Sign == "Plus", Pn * 2, (1 - Pn)*2 + 0.0002)) %>% 
    mutate(logP = -log10(P),
           P_score = if_else(Sign == "Plus", logP, -logP))
}

#run
Colony <- "0"; Comb <- 0; Value <- 0; Med <- 0; Pn <- 0
Sign <- "0"; P <- 0; logP <- 0; P_score <- 0
pv0 <- tibble(Colony, Comb, Value, Med, Pn, Sign, P, logP, P_score)
for(i in 1:5){
  colony <- colonyname[i]
  d <- DL_t_l %>% filter(Colony == colony) %>% mutate(Syn = L_val)
  pv <- pv_data(d)
  n <- ncol(pv)
  pv <- pv %>% mutate(Colony = colony) %>% select(n+1, 1:n)
  pv0 <- pv0 %>% bind_rows(pv)
}
pv_nloc <- pv0 %>% slice(-1)
```

##Proximity index data
```{r}
#example of frequency distribution in simulated value (Figure 5B)
DL_t_l_comb22 <- DL_t_l %>% filter(Colony == "LOC", Comb == 22)
DL_t_l_comb22 %>% filter(index != 1) %>% 
  group_by(Comb) %>% summarise(mean = mean(L_val), sd = sd(L_val))
#mean 16033.41	sd 1235.96
DL_t_l_comb22 %>% filter(index == 1) %>% pull(L_val)
#16739
ggplot(data = DL_t_l_comb22 %>% filter(index != 1)) + 
  geom_histogram(aes(x = L_val)) + 
  scale_y_continuous(expand = c(0,0)) + 
  theme_classic() + 
  theme(aspect.ratio = 1, axis.title = element_blank(), axis.text = element_blank())
ggsave("data/Figure/Figure5/Figure5B.pdf", width = 3, height = 3, units = "in")

#proximity index data (z_value = proximity index)
DL_t_l_sim_s <- DL_t_l %>% 
  group_by(Colony, Comb) %>% 
  filter(index != 1) %>% 
  summarise(Mean = mean(L_val), SD = sd(L_val))
zv_prox <- pv_nloc %>% 
  left_join(DL_t_l_sim_s, join_by(Colony, Comb)) %>% 
  mutate(z_value = (Value - Mean) / SD) %>% 
  left_join(ID_att_comb, join_by(Colony, Comb))
```

##Heatmap for proximity index (Figure 5C)
```{r}
#matrix function
zv_mat_data <- function(d, i){
  id_i <- idlist[i] %>% unlist()
  size <- length(id_i)
  ID1 <- rep(id_i, each = size)
  ID2 <- rep(id_i, size)
  z_value <- rep(NA, length(ID1))
  Dp0 <- tibble(ID1, ID2, z_value)
  ID1n <- rep(1:size, each = size)
  ID2n <- rep(1:size, size)
  Dp <- tibble(ID1, ID2, z_value, ID1n, ID2n) %>% 
    mutate(sum = ID1n + ID2n, dif = abs(ID1n - ID2n)) %>% 
    distinct(sum, dif, .keep_all = TRUE) %>% 
    filter(ID1 != ID2) %>% 
    select(ID1, ID2, z_value) %>% 
    mutate(z_value = d$z_value)
  Dpt <- Dp %>% select(ID1 = ID2, ID2 = ID1, z_value)
  Dp %>% bind_rows(Dpt) %>% bind_rows(Dp0) %>% 
    distinct(ID1, ID2, .keep_all = TRUE) %>% 
    mutate(ID1 = as.factor(ID1), ID2 = as.factor(ID2)) %>% 
    arrange(ID1, ID2) %>% 
    pivot_wider(names_from = ID2, values_from = z_value) %>% 
    column_to_rownames(var = "ID1")
}

#pheatmap function
zv_hm <- function(d, i, name, clu){
  colony <- colonyname[i]
  ID <- idlist[i] %>% unlist()
  Caste <- castelist[i] %>% unlist()
  anno <- tibble(ID, Caste) %>% column_to_rownames(var = "ID")
  Caste_col <- c("#AF528D","#54B345","grey")
  names(Caste_col) <- c("BF","BM","NB")
  anno_color <- list(Caste = Caste_col)
  RdYlBu <- c("#313695", "#4575b4", "#74add1", "#abd9e9", "#e0f3f8", "#ffffbf", 
              "#fee090", "#fdae61", "#f46d43", "#d73027","#a50026")
  
  val <- d %>% rowSums(1:length(ID)) %>% as_tibble() %>% pull(value)
  callback = function(hc, mat){
    dend = reorder(as.dendrogram(hc), wts = val)
    as.hclust(dend)
  }
  threshold <- 10
  
  pheatmap(
    mat = d,
    na_col = "white",
    annotation_row = anno, 
    annotation_col = anno,
    annotation_colors = anno_color,
    cellheight = 10, cellwidth = 10,
    border_color = "NA",
    breaks = c(seq(zv_min, threshold, length.out = 99), threshold + 1),
    color = colorRampPalette(RdYlBu)(n=100),
    annotation_legend = FALSE,
    annotation_names_row = name, annotation_names_col = name,
    show_rownames = name, show_colnames = name,
    legend = name,
    cluster_rows = clu, cluster_cols = clu,
    clustering_method = "ward.D2",
    clustering_callback = callback,
    filename = paste0("data/Figure/Figure5/Figure5C/", colony, "_name", name, "_clu", clu, ".pdf")
    )
}

#run
for(i in 1:5){
  colony <- colonyname[i]
  d <- zv_prox %>% filter(Colony == colony)
  zv_mat <- zv_mat_data(d, i)
  zv_hm(zv_mat, i, name=F, clu=T)
  zv_hm(zv_mat, i, name=T, clu=T)
}
```

##Heatmap for p-value (Figure 5D)
```{r}
#id sequence from figure 5C
id_1_prox <- c("M4", "M14", "2B34", "2B37", "2B42", "2B43", "2B45", "2B35", "2B33", "2B38", "2B49",
            "2B48", "2B51", "2B32", "2B39", "2B46", "2B41", "2B36", "2B44", "2B40", "2B47", "2B50")
id_2_prox <- c("2D11", "H11", "2D13", "2D33", "2D16", "2D35", "2D30", "2D31", "2D24",
               "2D34", "2D28", "2D26", "2D29", "2D19", "2D12", "2D23", "2D32")
id_3_prox <- c("G16", "H13", "HG19", "HG21", "HG14", "HG10", "HG16", "HG18", "HG11", "HG4",
               "HG23", "HG22", "HG24", "HG8", "HG5", "HG15", "HG12", "HG7", "HG13")
id_4_prox <- c("HT20", "HT16", "HT19", "R21", "HT17", "HT18", "HT3", "HT15", "HT21", "HT7", "HT14",
               "HT10", "HT11", "HT2", "HT13", "HT9", "HT12", "HT5", "HT1", "HT4", "HT6", "L63")
id_5_prox <- c("LOC19", "LOC22", "LOC7", "LOC8", "LOC5", "LOC23", "OC1", "L55", 
               "LOC20", "LOC14", "LOC9", "LOC17", "LOC13", "LOC16", "LOC6", 
               "LOC10", "LOC4", "LOC18", "LOC2", "LOC1", "LOC15", "LOC11")
idlist_prox <- list(id_1_prox, id_2_prox, id_3_prox, id_4_prox, id_5_prox)

#matrix function
pv_mat_data <- function(d, i, order){
  id_i <- idlist[i] %>% unlist()
  size <- length(id_i)
  ID1 <- rep(id_i, each = size)
  ID2 <- rep(id_i, size)
  P_score <- rep(NA, length(ID1))
  Dp0 <- tibble(ID1, ID2, P_score)
  ID1n <- rep(1:size, each = size)
  ID2n <- rep(1:size, size)
  Dp <- tibble(ID1, ID2, P_score, ID1n, ID2n) %>% 
    mutate(sum = ID1n + ID2n, dif = abs(ID1n - ID2n)) %>% 
    distinct(sum, dif, .keep_all = TRUE) %>% 
    filter(ID1 != ID2) %>% 
    select(ID1, ID2, P_score) %>% 
    mutate(P_score = d$P_score)
  Dpt <- Dp %>% select(ID1 = ID2, ID2 = ID1, P_score)
  Dp %>% bind_rows(Dpt) %>% bind_rows(Dp0) %>% 
    distinct(ID1, ID2, .keep_all = TRUE) %>% 
    mutate(ID1 = factor(ID1, levels = order),
           ID2 = factor(ID2, levels = order)) %>% 
    arrange(ID1, ID2) %>% 
    pivot_wider(names_from = ID2, values_from = P_score) %>% 
    column_to_rownames(var = "ID1")
}

#pheatmap function
pv_hm <- function(d, i, name, clu){
  colony <- colonyname[i]
  ID <- idlist[i] %>% unlist()
  Caste <- castelist[i] %>% unlist()
  anno <- tibble(ID, Caste) %>% column_to_rownames(var = "ID")
  Caste_col <- c("#AF528D","#54B345","grey")
  names(Caste_col) <- c("BF","BM","NB")
  anno_color <- list(Caste = Caste_col)
  RdYlBu <- c("#313695", "#4575b4", "#74add1", "#abd9e9", "#e0f3f8", "#ffffbf", 
              "#fee090", "#fdae61", "#f46d43", "#d73027","#a50026")
  pheatmap(
    mat = d,
    na_col = "white",
    annotation_row = anno, 
    annotation_col = anno,
    annotation_colors = anno_color,
    cellheight = 10, cellwidth = 10,
    border_color = "NA",
    breaks = seq(log10(0.0002), -log10(0.0002), length.out=100),
    color = colorRampPalette(RdYlBu)(n=100),
    annotation_legend = FALSE,
    annotation_names_row = name, annotation_names_col = name,
    show_rownames = name, show_colnames = name,
    legend = name,
    cluster_rows = clu, cluster_cols = clu,
    filename = paste0("data/Figure/Figure5/Figure5D/", colony, "_name", name, "_clu", clu, ".pdf")
    )
}

#run
for(i in 1:5){
  colony <- colonyname[i]
  d <- zv_prox %>% filter(Colony == colony)
  order <- idlist_prox[i] %>% unlist()
  pv_mat <- pv_mat_data(d, i, order)
  pv_hm(pv_mat, i, varname, name=F, clu=F)
}
```


##Network (Figure 5E)
```{r}
#network function
net_vis_z <- function(d, i, var){
  colony <- colonyname[i]
  ID <- idlist[i] %>% unlist()
  size <- length(ID)
  caste <- castelist[i] %>% unlist()
  
  d_syn <- d %>% filter(Colony == colony) %>% 
    filter(Sign == "Plus", P < 0.05) %>% select(ID1, ID2, !!sym(var))
  
  id_data <- tibble(ID, caste) %>% 
    mutate(caste_ind = case_when(caste == "BF" ~ 1, caste == "BM" ~ 2, caste == "NB" ~ 3))
  
  set.seed(15)
  pdf(paste0("data/Figure/Figure5/Figure5E/", colony, ".pdf"), 
    width = 3, height = 3)
  net_ss <- graph.data.frame(d = d_syn, vertices = id_data, directed = FALSE)
  E(net_ss)$width <- d_syn %>% pull(!!sym(var)) * 0.1
  col <- c("#AF528D","#54B345","grey")
  V(net_ss)$color <- col[V(net_ss)$caste_ind]
  w <- d_syn %>% pull(!!sym(var))
  c = (1 - ((w-min(w)) / (max(w)-min(w)))) * 0.8
  plot(net_ss, 
       edge.color = rgb(c,c,c), 
       vertex.label = NA, 
       vertex.frame.color = NA,
       vertex.label.color = "black", 
       vertex.label.cex = 0.1, 
       vertex.size = 10)
  dev.off()
}

#run
for(i in 1:5){
  var <- "z_value"
  net_vis_z(zv_prox, i, var)
}
```

##Network & behavioral cluster (Figure S5D)
```{r}
#network function
net_vis_z_behtype <- function(d, i, var){
  colony <- colonyname[i]
  ID <- idlist[i] %>% unlist()
  size <- length(ID)
  cluster <- d_behtype_mean_d %>% filter(Colony == colony) %>% pull(Cluster)
  
  d_syn <- d %>% filter(Colony == colony) %>% 
    filter(Sign == "Plus", P < 0.05) %>% select(ID1, ID2, !!sym(var))
  
  id_data <- tibble(ID, cluster) %>% 
    mutate(cluster_ind = case_when(
      cluster == "1" ~ 1, cluster == "2" ~ 2, cluster == "3" ~ 3, cluster == "4" ~ 4, 
      cluster == "5" ~ 5, cluster == "6" ~ 6, cluster == "7" ~ 7, cluster == "8" ~ 8
      ))
  
  set.seed(15)
  pdf(paste0("data/Figure/Figure5/FigureS5D/", colony, ".pdf"), 
    width = 3, height = 3)
  net_ss <- graph.data.frame(d = d_syn, vertices = id_data, directed = FALSE)
  E(net_ss)$width <- d_syn %>% pull(!!sym(var)) * 0.1
  V(net_ss)$color <- colorlist[V(net_ss)$cluster_ind]
  w <- d_syn %>% pull(!!sym(var))
  c = (1 - ((w-min(w)) / (max(w)-min(w)))) * 0.8
  plot(net_ss, 
       edge.color = rgb(c,c,c), 
       vertex.label = NA, vertex.frame.color = NA, vertex.label.color = "black", 
       vertex.label.cex = 0.1, vertex.size = 10)
  dev.off()
}

#run
for(i in 1:5){
  var <- "z_value"
  net_vis_z_behtype(zv_prox, i, var)
}
```


#3 Statistical analysis using synchrony index

##Add individual attribute
```{r}
zv_prox_att <- zv_prox %>% 
  left_join(ID_att_1, by = join_by(Colony, ID1)) %>% 
  left_join(ID_att_2, by = join_by(Colony, ID2)) %>% 
  mutate(Sig = case_when(
    Sign == "Plus" & P < 0.05 ~ "Syn", Sign == "Minus" & P < 0.05 ~ "Asyn", TRUE ~ "NS")
    ) %>% 
  mutate(Sig = factor(Sig, levels = c("Syn", "NS", "Asyn"))) %>% 
  select(Colony, ID1, ID2, Comb, z_value, Sign, P, Sig,
         Caste1, Caste2, Sex1, Sex2, AgeD1, AgeD2, Weight1, Weight2)

zv_prox_att_t <- zv_prox_att %>% 
  select(Colony, ID1 = ID2, ID2 = ID1, Comb, z_value, Sign, P, Sig,
         Caste1 = Caste2, Caste2 = Caste1, Sex1 = Sex2, Sex2 = Sex1, 
         AgeD1 = AgeD2, AgeD2 = AgeD1, Weight1 = Weight2, Weight2 = Weight1)
zv_prox_att_all <- zv_prox_att %>% bind_rows(zv_prox_att_t) %>% arrange(Colony, Comb)
```

##Colony-level (Figures S5A & 5F)
```{r}
#significant/ns proportion per colony (figure S5A)
zv_prox_att_prop <- zv_prox_att %>% 
  group_by(Colony, Sig) %>% summarise(n = n()) %>% mutate(p = n / sum(n))

ggplot(data = zv_prox_att_prop, aes(x = Colony, y = p, fill = Sig)) + 
  geom_bar(stat = "identity", position = "stack", width = 0.7) + 
  scale_fill_manual(values = c("#E6855E", "#BEBEBE", "#44A5CB")) + 
  scale_y_continuous(expand = c(0,0)) + 
  theme_classic() + 
  theme(aspect.ratio = 1.5, legend.position = "none",
        axis.title = element_blank(), axis.text = element_blank())
ggsave("data/Figure/Figure5/FigureS5A.pdf", width = 3, height = 4.5, units = "in")

#model
lm_prox_colony <- lm(data = zv_prox_att %>% mutate(Colony = as.factor(Colony)), 
                     formula = z_value ~ Colony)
summary(lm_prox_colony)
#Multiple R-squared:  0.1307,	Adjusted R-squared:  0.1272 

#deviation of EMMs from shuffling control (0)
emm <- emmeans(lm_prox_colony, ~ Colony)
test_res <- test(emm, null = 0) %>% as_tibble()
write.csv(test_res, row.names = F, "data/statistical_result/Figure5F.csv")

#visualization (figure 5F)
p_lm_prox_colony <- ggpredict(lm_prox_colony, terms = c("Colony [all]"))
ggplot() + 
  geom_hline(yintercept = 0, linetype = "dashed", color = "grey") + 
  geom_beeswarm(data = zv_prox_att, aes(x = Colony, y = z_value, color = Sig), 
                size = 0.8, cex = 1, alpha = 0.75) + 
  geom_errorbar(data = p_lm_prox_colony, 
                aes(x = x, ymax = predicted, ymin = predicted),
                linewidth = 1, width = 0.5) + 
  scale_color_manual(values = c("#F57446", "#595757", "#669DC8")) + 
  theme_classic() + 
  theme(aspect.ratio = 1.25, legend.position = "none",
        axis.title = element_blank(), axis.text = element_blank())
ggsave("data/Figure/Figure5/Figure5F.pdf", width = 4, height = 5, units = "in")
```

##Caste combination (Figure 5G)
```{r}
#data
zv_prox_att_castecomb <- zv_prox_att %>% 
  unite("Caste_comb", c(Caste1, Caste2), sep = "_", remove = F) %>% 
  mutate(Caste_comb = if_else(Caste_comb == "N_B", "B_N", Caste_comb))

#model
lmm_prox_castecomb <- lmerTest::lmer(
  data = zv_prox_att_castecomb, formula = z_value ~ Caste_comb + (1|Colony)
)
summary(lmm_prox_castecomb)
r.squaredGLMM(lmm_prox_castecomb)
#R2m 0.1536915 R2c 0.296256
Anova(lmm_prox_castecomb, test.statistic = "Chisq")
##x2=217.75, Df=2, p<0.001
emm_res <- emmeans(lmm_prox_castecomb, pairwise ~ Caste_comb)
pair_res <- pairs(emm_res, adjust = "holm") %>% as_tibble()
write.csv(pair_res, row.names = F, "data/statistical_result/Figure5G.csv")

#visualization
p_lmm_prox_castecomb <- ggpredict(lmm_prox_castecomb, terms = c("Caste_comb [all]"))
ggplot() + 
  geom_beeswarm(data = zv_prox_att_castecomb, aes(x = Caste_comb, y = z_value), 
                size = 0.8, cex = 0.95, alpha = 0.4) + 
  geom_errorbar(data = p_lmm_prox_castecomb, 
                aes(x = x, ymax = predicted, ymin = predicted),
                size = 1, width = 0.5) + 
  theme_classic() + 
  theme(aspect.ratio = 1.5, legend.position = "none",
        axis.title = element_blank(), axis.text = element_blank())
ggsave("data/Figure/Figure5/Figure5G.pdf", width = 4, height = 6, units = "in")
```

##Sex combination (Figures S5B & S5C)
```{r}
#data
zv_prox_att_sexcomb <- zv_prox_att %>% 
  unite("Sex_comb", c(Sex1, Sex2), sep = "_", remove = F) %>% 
  mutate(Sex_comb = case_when(
    Sex_comb == "F_BF" ~ "BF_F", 
    Sex_comb == "M_BF" ~ "BF_M", 
    Sex_comb == "F_BM" ~ "BM_F", 
    Sex_comb == "M_BM" ~ "BM_M", 
    Sex_comb == "M_F" ~ "F_M", 
    TRUE ~ Sex_comb
    )) %>% 
  filter(Sex_comb != "BF_BM", Sex_comb != "BM_BF")


#B-NB data
zv_prox_att_sexcomb_bnb <- zv_prox_att_sexcomb %>% filter(Caste1 == "B", Caste2 == "N")
lmm_prox_sexcomb_bnb <- lmerTest::lmer(
  data = zv_prox_att_sexcomb_bnb, 
  formula = z_value ~ Sex_comb + (1|Colony)
)
r.squaredGLMM(lmm_prox_sexcomb_bnb)
#R2m 0.05494589 R2c 0.5168899
Anova(lmm_prox_sexcomb_bnb, test.statistic = "Chisq")
##x2=20.76, Df=3, p<0.001
emm_res <- emmeans(lmm_prox_sexcomb_bnb, pairwise ~ Sex_comb)
pair_res <- pairs(emm_res, adjust = "holm") %>% as_tibble()
write.csv(pair_res, row.names = F, "data/statistical_result/FigureS5B.csv")

lmm_prox_sexcomb_bnb_p <- ggpredict(lmm_prox_sexcomb_bnb, terms = "Sex_comb")
ggplot() + 
  geom_beeswarm(data = zv_prox_att_sexcomb_bnb, 
                aes(x = Sex_comb, y = z_value), 
                size = 1.5, cex = 2.5, alpha = 0.4) + 
  geom_errorbar(data = lmm_prox_sexcomb_bnb_p, 
                aes(x = x, ymax = predicted, ymin = predicted),
                size = 1, width = 0.5) + 
  theme_classic() + 
  theme(aspect.ratio = 1.65, legend.position = "none",
        axis.title = element_blank(), axis.text = element_blank())
ggsave("data/Figure/Figure5/FigureS5B.pdf", width = 2.5, height = 4, units = "in")


#NB-NB data
zv_prox_att_sexcomb_nbnb <- zv_prox_att_sexcomb %>% filter(Caste1 == "N", Caste2 == "N")
lmm_prox_sexcomb_nbnb <- lmerTest::lmer(
  data = zv_prox_att_sexcomb_nbnb, 
  formula = z_value ~ Sex_comb + (1|Colony)
)
r.squaredGLMM(lmm_prox_sexcomb_nbnb)
#R2m 0.001990926 R2c 0.1697721
Anova(lmm_prox_sexcomb_nbnb, test.statistic = "Chisq")
##x2=1.8604, Df=2, p=0.395
emm_res <- emmeans(lmm_prox_sexcomb_nbnb, pairwise ~ Sex_comb)
pair_res <- pairs(emm_res, adjust = "holm") %>% as_tibble()
write.csv(pair_res, row.names = F, "data/statistical_result/FigureS5C.csv")

lmm_prox_sexcomb_nbnb_p <- ggpredict(lmm_prox_sexcomb_nbnb, terms = "Sex_comb")
ggplot() + 
  geom_beeswarm(data = zv_prox_att_sexcomb_nbnb, 
                aes(x = Sex_comb, y = z_value), 
                size = 1, cex = 1.5, alpha = 0.4) + 
  geom_errorbar(data = lmm_prox_sexcomb_nbnb_p, 
                aes(x = x, ymax = predicted, ymin = predicted),
                size = 1, width = 0.5) + 
  theme_classic() + 
  theme(aspect.ratio = 1.65, legend.position = "none",
        axis.title = element_blank(), axis.text = element_blank())
ggsave("data/Figure/Figure5/FigureS5C.pdf", width = 2.5, height = 4, units = "in")
```

##Behavioral cluster combination (Figure 5H)
```{r}
#data
d_behtype_mean <- read.csv("data/processed_data/representative_cluster.csv")
d_behtype_mean1 <- d_behtype_mean %>% select(Colony, ID1 = ID, Cluster1 = Cluster)
d_behtype_mean2 <- d_behtype_mean %>% select(Colony, ID2 = ID, Cluster2 = Cluster)

zv_prox_att_behtype_comb <- zv_prox_att %>% 
  left_join(d_behtype_mean1, by = join_by(Colony, ID1)) %>% 
  left_join(d_behtype_mean2, by = join_by(Colony, ID2)) %>% 
  drop_na(everything()) %>% 
  unite("Cluster_comb", c(Cluster1, Cluster2), sep = "_", remove = F) %>% 
  mutate(Cluster_comb = case_when(
    Cluster_comb == "2_1" ~ "1_2", Cluster_comb == "3_1" ~ "1_3", Cluster_comb == "4_1" ~ "1_4", 
    Cluster_comb == "5_1" ~ "1_5", Cluster_comb == "6_1" ~ "1_6", Cluster_comb == "7_1" ~ "1_7", 
    Cluster_comb == "3_2" ~ "2_3", Cluster_comb == "4_2" ~ "2_4", Cluster_comb == "5_2" ~ "2_5", 
    Cluster_comb == "6_2" ~ "2_6", Cluster_comb == "7_2" ~ "2_7", 
    Cluster_comb == "4_3" ~ "3_4", Cluster_comb == "5_3" ~ "3_5", Cluster_comb == "6_3" ~ "3_6", 
    Cluster_comb == "7_3" ~ "3_7", 
    Cluster_comb == "5_4" ~ "4_5", Cluster_comb == "6_4" ~ "4_6", Cluster_comb == "7_4" ~ "4_7", 
    Cluster_comb == "6_5" ~ "5_6", Cluster_comb == "7_5" ~ "5_7", 
    Cluster_comb == "7_6" ~ "6_7", 
    TRUE ~ Cluster_comb
    )) %>% 
  mutate(Cluster_comb = as.factor(Cluster_comb))

#model
lmm_prox <- lmerTest::lmer(data = zv_prox_att_behtype_comb, 
                           formula = z_value ~ 1 + (1|Colony))
prox_mean <- fixef(lmm_prox)["(Intercept)"]

lmm_prox_behtypecomb <- lmerTest::lmer(
  data = zv_prox_att_behtype_comb, 
  formula = z_value ~ Cluster_comb + (1|Colony)
)
r.squaredGLMM(lmm_prox_behtypecomb)
#R2m 0.2560869 R2c 0.3941174
Anova(lmm_prox_behtypecomb, test.statistic = "Chisq")
##x2=411.91, Df=27, p<0.001

#deviation of EMMs from mean (syntime_chance)
emm <- emmeans(lmm_prox_behtypecomb, ~ Cluster_comb)
test_res <- test(emm, null = prox_mean) %>% as_tibble()
write.csv(test_res, row.names = F, "data/statistical_result/Figure5H.csv")

#visualization
zv_prox_att_all_behtype_comb <- zv_prox_att_all %>% 
  left_join(d_behtype_mean1, by = join_by(Colony, ID1)) %>% 
  left_join(d_behtype_mean2, by = join_by(Colony, ID2)) %>% 
  drop_na(everything()) %>% 
  mutate(Cluster1 = as.factor(Cluster1), Cluster2 = as.factor(Cluster2))
lmm_prox_all_behtypecomb <- lmerTest::lmer(
  data = zv_prox_att_all_behtype_comb, 
  formula = z_value ~ Cluster1 * Cluster2 + (1|Colony)
)
lmm_prox_all_behtypecomb_p <- ggpredict(lmm_prox_all_behtypecomb, terms = c("Cluster1 [all]", "Cluster2 [all]"))
colorlist <- c("#BE5378", "#CE7754", "#DFC44E", "#54B345", "#3B98B2", "#5F7EB2", "#808080")
ggplot() + 
  geom_hline(yintercept = prox_mean, linetype = "dashed", color = "grey") + 
  geom_violin(data = zv_prox_att_all_behtype_comb, 
              aes(x = Cluster1, y = z_value, color = Cluster2),
              position = position_dodge(width = 0.8), width = 0.5, scale = "width") + 
  geom_errorbar(data = lmm_prox_all_behtypecomb_p, 
                aes(x = x, ymax = predicted, ymin = predicted, color = group), 
                position = position_dodge(width = 0.8), size = 1, width = 0.5) + 
  scale_color_manual(values = colorlist) + 
  theme_classic() + 
  theme(aspect.ratio = 0.25, legend.position = "none",
        axis.title = element_blank(), axis.text = element_blank())
ggsave("data/Figure/Figure5/Figure5H.pdf", width = 16, height = 4, units = "in")
```


#4 Correlation between synchrony and proximity index (Figure 5I)
```{r}
zv_sync_prox <- zv_syntime %>% 
  select(Colony, Comb, Synchrony = z_value) %>% 
  left_join(zv_prox %>% select(Colony, Comb, Proximity = z_value), join_by(Colony, Comb))

#Pearson correlation test
cor.test(data = zv_sync_prox, ~ Synchrony + Proximity)
#t=17.377 df=998 r=0.482 p<0.001

#plot
ggplot(zv_sync_prox) + 
  geom_point(aes(x = Synchrony, y = Proximity), size = 2, alpha = 0.2) + 
  theme_classic() + 
  theme(aspect.ratio = 1, 
        axis.title = element_blank(), axis.text = element_blank())
ggsave("data/Figure/Figure5/Figure5I.pdf", width = 4, height = 4, units = "in")

#per colony
cor.test(data = zv_sync_prox %>% filter(Colony == "2B"), ~ Synchrony + Proximity)
#t=7.5211 df=229 r=0.4450675 p<0.001
cor.test(data = zv_sync_prox %>% filter(Colony == "2D"), ~ Synchrony + Proximity)
#t=6.9187 df=134 r=0.5130338 p<0.001
cor.test(data = zv_sync_prox %>% filter(Colony == "HG"), ~ Synchrony + Proximity)
#t=4.5117 df=169 r=0.3278704 p<0.001
cor.test(data = zv_sync_prox %>% filter(Colony == "HT"), ~ Synchrony + Proximity)
#t=6.7378 df=229 r=0.4067506 p<0.001
cor.test(data = zv_sync_prox %>% filter(Colony == "LOC"), ~ Synchrony + Proximity)
#t=8.2015 df=229 r=0.4764874 p<0.001
```

