---
title: "Figure 4"
output: html_notebook
---

#1 Permutation 10000 times for each dyad

## 30-day rest/active vector for each id function
```{r}
vector_const_sleact <- function(k){
  
  colony <- colonyname[k]
  id <- idlist[[k]]
  size <- length(id)
  dataset <- dayname[[k]]
  
  #prepare initial variable for each id
  for(i in 1:size){
    assign(paste0("d_",i), "0", envir = .GlobalEnv)
  }
  
  #30-day vector for each id
  for(name in dataset){
    d <- fread(paste0("data/processed_data/locsleep_data/", colony, "/", name, ".csv")) %>% 
      mutate(Time10s = ceiling((Time-0.01)/10)) %>% 
      distinct(Time10s, .keep_all = TRUE) %>% 
      select(-Time, -Time10s)
    for(i in 1:size){
      di <- d %>% pull(i) %>% replace_na("0")
      nth <- paste0("d_",i)
      assign(nth, c(get(nth), di), envir = .GlobalEnv)
    }
  }
  for(i in 1:size){
    nth <- paste0("d_",i)
    assign(nth, get(nth)[-1], envir = .GlobalEnv)
    di <- get(paste0("d_",i)) %>% tibble() %>% 
      mutate(Sleep = if_else(. == "S", 1, 0)) %>% pull(Sleep)
    assign(nth, di, envir = .GlobalEnv)
  }
}
```

##Permutation for rest or active function
```{r}
##synchronized times calculation function (rest + active)
syn_time <- function(k, ran, ID1, ID2){
  if(ran[k] == 1){ID1r <- ID1}
  if(ran[k] != 1){ID1r <- ID1[c(ran[k]:N,1:(ran[k]-1))]}
  ID12 <- ID1r + ID2
  Dt <<- sum(ID12 == 2 | ID12 == 0)
}

#permutation (circular) test for rest/active function
syn_time_pt.test <- function(ran, ID1, ID2, nperm){
  DT <- 0
  for(k in 1:(nperm+1)){
    syn_time(k, ran, ID1, ID2)
    DT <- c(DT, Dt)
  }
  DT <<- DT[-1]
}

#permutation test for rest/active performing function
syn_time_pttest <- function(colony, nperm){
  
  size <- idlist[colony] %>% unlist() %>% length()
  pb <- txtProgressBar(min = 1, max = size*(size-1)/2, style = 3)
  DTs <- list(0)
  N <- length(d_1)
  
  #determine sequence to permute vector randomly
  set.seed(10)
  ran <- sample(2:N, nperm, replace = FALSE)
  ran <- c(1, ran) #add actual value to the first element
  
  for(i in 1:(size-1)){
    #30-day vector for ID1
    ID1 <- get(paste0("d_",i))
    
    for(j in (i+1):size){
      #progress bar
      n <- size*(size-1)/2 - (size-i+1)*(size-i)/2 + (j-i)
      setTxtProgressBar(pb, n)
      #30-day vector for ID2
      ID2 <- get(paste0("d_",j))
      #permutation test
      syn_sleact_pt.test(ran, ID1, ID2, nperm)
      DTs[n] <- list(DT)
    }
  }
  DT_s <<- DTs
}
```

##Run function
```{r}
nperm <- 9999

for(l in 1:5){
  vector_const_sleact(l)
  system.time(syn_time_pttest(l, nperm))
  
  size <- idlist[l] %>% unlist() %>% length()
  DT_s <- data.frame(DT_s)
  colnames(DT_s) <- 1:(size*(size-1)/2)
  nth <- paste0("DT_", l)
  assign(nth, DT_s)
}
#output data: DT_1~5


#trial for the first dyad
#colony <- 1
#size <- idlist[colony] %>% unlist() %>% length()
#DTs <- list(0)
#N <- length(d_1)
#set.seed(10)
#ran <- sample(2:N, nperm, replace = FALSE)
#ran <- c(1, ran)
#ID1 <- get(paste0("d_",1))
#ID2 <- get(paste0("d_",2))
#syn_time_pt.test(ran, ID1, ID2, nperm)
#n <- size*(size-1)/2 - (size-1+1)*(size-1)/2 + (2-1)
#DTs[n] <- list(DT)
#DT_s <- DTs[[1]] %>% as.vector()
#DT_s <- tibble(DT_s)
```

##Integrate data
```{r}
trans_tidy_time <- function(d, i){
  colony <- colonyname[i]
  d %>% 
    mutate(index = 1:10000) %>% 
    pivot_longer(cols = -index, names_to = "Comb", values_to = "T_val") %>% 
    mutate(Comb = as.integer(Comb)) %>% 
    arrange(Comb) %>% 
    mutate(Colony = colony) %>% 
    select(Colony, Comb, index, T_val)
}

DT_1_l <- trans_tidy_time(DT_1, 1)
DT_2_l <- trans_tidy_time(DT_2, 2)
DT_3_l <- trans_tidy_time(DT_3, 3)
DT_4_l <- trans_tidy_time(DT_4, 4)
DT_5_l <- trans_tidy_time(DT_5, 5)

DT_t_l <- bind_rows(DT_1_l, DT_2_l, DT_3_l, DT_4_l, DT_5_l)
```

##Dyad information
```{r}
ID_att <- read.csv("data/pre_info_data/individual_information.csv") %>% mutate(Birth = as.Date(Birth))

ID_att_1 <- ID_att %>% select(Colony, ID1 = ID, Caste1 = Caste, Sex1 = Sex, AgeD1 = AgeD, Weight1 = Weight)
ID_att_2 <- ID_att %>% select(Colony, ID2 = ID, Caste2 = Caste, Sex2 = Sex, AgeD2 = AgeD, Weight2 = Weight)

idattcomb_each <- function(colony, size){
  rep(ID_att %>% filter(Colony == colony) %>% pull(ID), each = size)
}
idattcomb <- function(colony, size){
  rep(ID_att %>% filter(Colony == colony) %>% pull(ID), size)
}

ID1 <- "0"; ID2 <- "0"; Colony <- "0"; ID1n <- 0; ID2n <- 0 
for(i in 1:5){
  colony <- colonyname[i]
  size <- idlist[i] %>% unlist() %>% length()
  
  ID1_i <- idattcomb_each(colony, size); ID2_i <- idattcomb(colony, size)
  Colony_i <- rep(colony, size^2)
  ID1n_i <- rep(1:size, each = size); ID2n_i <- rep(1:size, size)
  
  ID1 <- c(ID1, ID1_i); ID2 <- c(ID2, ID2_i)
  Colony <- c(Colony, Colony_i)
  ID1n <- c(ID1n, ID1n_i); ID2n <- c(ID2n, ID2n_i)
}
Comb <- c(1:231, 1:136, 1:171, 1:231, 1:231)

ID_att_comb <- tibble(Colony, ID1, ID2, ID1n, ID2n) %>% 
  slice(-1) %>% filter(ID1 != ID2) %>% 
  mutate(sum = ID1n + ID2n, dif = abs(ID1n - ID2n)) %>% 
  group_by(Colony) %>% distinct(sum, dif, .keep_all = T) %>% 
  ungroup() %>% mutate(Comb = Comb) %>% 
  select(Colony, Comb, ID1, ID2)
```


#2 Visualization of p-value & synchrony index

##p-value data
```{r}
#p-value summarize function
pv_data <- function(d){
  d %>% group_by(Comb) %>% 
    summarise(Value = Syn[index == 1],
              Med = median(Syn),
              Pn = sum(Syn >= Syn[index == 1]) / 10000) %>% 
    mutate(Sign = if_else(Pn < 0.5, "Plus", "Minus")) %>% 
    mutate(P = if_else(Sign == "Plus", Pn * 2, (1 - Pn)*2 + 0.0002)) %>% 
    mutate(logP = -log10(P),
           P_score = if_else(Sign == "Plus", logP, -logP))
}

#run
Colony <- "0"; Comb <- 0; Value <- 0; Med <- 0; Pn <- 0
Sign <- "0"; P <- 0; logP <- 0; P_score <- 0
pv0 <- tibble(Colony, Comb, Value, Med, Pn, Sign, P, logP, P_score)
for(i in 1:5){
  colony <- colonyname[i]
  d <- DT_t_l %>% filter(Colony == colony) %>% mutate(Syn = T_val)
  pv <- pv_data(d)
  n <- ncol(pv)
  pv <- pv %>% mutate(Colony = colony) %>% select(n+1, 1:n)
  pv0 <- pv0 %>% bind_rows(pv)
}
pv_syntime <- pv0 %>% slice(-1)
```

##Synchrony index data
```{r}
#example of frequency distribution in simulated value (Figure 4B)
DT_t_l_comb22 <- DT_t_l %>% filter(Colony == "LOC", Comb == 22) 

DT_t_l_comb22 %>% filter(index != 1) %>% 
  group_by(Comb) %>% summarise(mean = mean(T_val), sd = sd(T_val))
##mean 132048.1	sd 3805.511
DT_t_l_comb22 %>% filter(index == 1) %>% pull(T_val)
##140861
ggplot(data = DT_t_l_comb22 %>% filter(index != 1)) + 
  geom_histogram(aes(x = T_val)) + 
  scale_y_continuous(expand = c(0,0)) + 
  theme_classic() + 
  theme(aspect.ratio = 1, axis.title = element_blank(), axis.text = element_blank())
ggsave("data/Figure/Figure4/Figure4B.pdf", width = 3, height = 3, units = "in")

#synchrony index data (z_value = synchrony index)
DT_t_l_sim_s <- DT_t_l %>% 
  group_by(Colony, Comb) %>% 
  filter(index != 1) %>% 
  summarise(Mean = mean(T_val), SD = sd(T_val))
zv_syntime <- pv_syntime %>% 
  left_join(DT_t_l_sim_s, join_by(Colony, Comb)) %>% 
  mutate(z_value = (Value - Mean) / SD) %>% 
  left_join(ID_att_comb, join_by(Colony, Comb))
```

##Heatmap for synchrony index (Figure 4C)
```{r}
#matrix function
zv_mat_data <- function(d, i){
  id_i <- idlist[i] %>% unlist()
  size <- length(id_i)
  ID1 <- rep(id_i, each = size)
  ID2 <- rep(id_i, size)
  z_value <- rep(NA, length(ID1))
  Dp0 <- tibble(ID1, ID2, z_value)
  ID1n <- rep(1:size, each = size)
  ID2n <- rep(1:size, size)
  Dp <- tibble(ID1, ID2, z_value, ID1n, ID2n) %>% 
    mutate(sum = ID1n + ID2n, dif = abs(ID1n - ID2n)) %>% 
    distinct(sum, dif, .keep_all = TRUE) %>% 
    filter(ID1 != ID2) %>% 
    select(ID1, ID2, z_value) %>% 
    mutate(z_value = d$z_value)
  Dpt <- Dp %>% select(ID1 = ID2, ID2 = ID1, z_value)
  Dp %>% bind_rows(Dpt) %>% bind_rows(Dp0) %>% 
    distinct(ID1, ID2, .keep_all = TRUE) %>% 
    mutate(ID1 = as.factor(ID1), ID2 = as.factor(ID2)) %>% 
    arrange(ID1, ID2) %>% 
    pivot_wider(names_from = ID2, values_from = z_value) %>% 
    column_to_rownames(var = "ID1")
}

#pheatmap function
zv_hm <- function(d, i, name, clu){
  colony <- colonyname[i]
  ID <- idlist[i] %>% unlist()
  Caste <- castelist[i] %>% unlist()
  anno <- tibble(ID, Caste) %>% column_to_rownames(var = "ID")
  Caste_col <- c("#AF528D","#54B345","grey")
  names(Caste_col) <- c("BF","BM","NB")
  anno_color <- list(Caste = Caste_col)
  RdYlBu <- c("#313695", "#4575b4", "#74add1", "#abd9e9", "#e0f3f8", "#ffffbf", 
              "#fee090", "#fdae61", "#f46d43", "#d73027","#a50026")
  
  val <- d %>% rowSums(1:length(ID)) %>% as_tibble() %>% pull(value)
  callback = function(hc, mat){
    dend = reorder(as.dendrogram(hc), wts = val)
    as.hclust(dend)
  }
  threshold <- 9.5
  
  pheatmap(
    mat = d,
    na_col = "white",
    annotation_row = anno, 
    annotation_col = anno,
    annotation_colors = anno_color,
    cellheight = 10, cellwidth = 10,
    border_color = "NA",
    breaks = c(seq(zv_min, threshold, length.out = 99), threshold + 1),
    color = colorRampPalette(RdYlBu)(n=100),
    annotation_legend = FALSE,
    annotation_names_row = name, annotation_names_col = name,
    show_rownames = name, show_colnames = name,
    legend = name,
    cluster_rows = clu, cluster_cols = clu,
    clustering_method = "ward.D2",
    clustering_callback = callback,
    filename = paste0("data/Figure/Figure4/Figure4C/", colony, "_name", name, "_clu", clu, ".pdf")
    )
}

#run
for(i in 1:5){
  colony <- colonyname[i]
  d <- zv_syntime %>% filter(Colony == colony)
  order <- idlist_syntime[i] %>% unlist()
  zv_mat <- zv_mat_data(d, i, order)
  zv_hm(zv_mat, i, varname, name=F, clu=T)
  zv_hm(zv_mat, i, varname, name=T, clu=T)
}
```


##Heatmap for p-value (Figure 4D)
```{r}
#id sequence from figure 4C
id_1_sync <- c("2B36", "M14", "M4", "2B33", "2B34", "2B35", "2B42", "2B32", "2B50", "2B39", "2B48", 
                "2B51", "2B37", "2B38", "2B43", "2B45", "2B47", "2B40", "2B49", "2B44", "2B41", "2B46")
id_2_sync <- c("2D19", "2D11", "H11", "2D13", "2D33", "2D30", "2D16", "2D24", "2D32", 
                    "2D23", "2D28", "2D35", "2D34", "2D26", "2D29", "2D12", "2D31")
id_3_sync <- c("HG10", "HG22", "HG18", "HG23", "HG19", "HG21", "HG14", "HG11", "HG12",
                    "HG8", "H13", "G16", "HG13", "HG7", "HG5", "HG4", "HG24", "HG16", "HG15")
id_4_sync <- c("HT21", "HT20", "HT15", "HT19", "HT18", "L63", "HT2", "HT9", "HT14", "HT1", "HT3",
                    "HT5", "HT4", "R21", "HT10", "HT7", "HT13", "HT16", "HT12", "HT6", "HT17", "HT11")
id_5_sync <- c("OC1", "L55", "LOC10", "LOC4", "LOC18", "LOC11", "LOC16", "LOC9",
                    "LOC13", "LOC2", "LOC15", "LOC6", "LOC17", "LOC23", "LOC20", 
                    "LOC7", "LOC5", "LOC1", "LOC22", "LOC8", "LOC14", "LOC19")
idlist_sync <- list(id_1_sync, id_2_sync, id_3_sync, id_4_sync, id_5_sync)

#matrix function
pv_mat_data <- function(d, i, order){
  id_i <- idlist[i] %>% unlist()
  size <- length(id_i)
  ID1 <- rep(id_i, each = size)
  ID2 <- rep(id_i, size)
  P_score <- rep(NA, length(ID1))
  Dp0 <- tibble(ID1, ID2, P_score)
  ID1n <- rep(1:size, each = size)
  ID2n <- rep(1:size, size)
  Dp <- tibble(ID1, ID2, P_score, ID1n, ID2n) %>% 
    mutate(sum = ID1n + ID2n, dif = abs(ID1n - ID2n)) %>% 
    distinct(sum, dif, .keep_all = TRUE) %>% 
    filter(ID1 != ID2) %>% 
    select(ID1, ID2, P_score) %>% 
    mutate(P_score = d$P_score)
  Dpt <- Dp %>% select(ID1 = ID2, ID2 = ID1, P_score)
  Dp %>% bind_rows(Dpt) %>% bind_rows(Dp0) %>% 
    distinct(ID1, ID2, .keep_all = TRUE) %>% 
    mutate(ID1 = factor(ID1, levels = order),
           ID2 = factor(ID2, levels = order)) %>% 
    arrange(ID1, ID2) %>% 
    pivot_wider(names_from = ID2, values_from = P_score) %>% 
    column_to_rownames(var = "ID1")
}

#pheatmap function
pv_hm <- function(d, i, name, clu){
  colony <- colonyname[i]
  ID <- idlist[i] %>% unlist()
  Caste <- castelist[i] %>% unlist()
  anno <- tibble(ID, Caste) %>% column_to_rownames(var = "ID")
  Caste_col <- c("#AF528D","#54B345","grey")
  names(Caste_col) <- c("BF","BM","NB")
  anno_color <- list(Caste = Caste_col)
  RdYlBu <- c("#313695", "#4575b4", "#74add1", "#abd9e9", "#e0f3f8", "#ffffbf", 
              "#fee090", "#fdae61", "#f46d43", "#d73027","#a50026")
  pheatmap(
    mat = d,
    na_col = "white",
    annotation_row = anno, 
    annotation_col = anno,
    annotation_colors = anno_color,
    cellheight = 10, cellwidth = 10,
    border_color = "NA",
    breaks = seq(log10(0.0002), -log10(0.0002), length.out=100),
    color = colorRampPalette(RdYlBu)(n=100),
    annotation_legend = FALSE,
    annotation_names_row = name, annotation_names_col = name,
    show_rownames = name, show_colnames = name,
    legend = name,
    cluster_rows = clu, cluster_cols = clu,
    filename = paste0("data/Figure/Figure4/", colony, "_name", name, "_clu", clu, ".pdf")
    )
}

#run
for(i in 1:5){
  colony <- colonyname[i]
  d <- zv_syntime %>% filter(Colony == colony)
  order <- idlist_sync[i] %>% unlist()
  pv_mat <- pv_mat_data(d, i, order)
  pv_hm(pv_mat, i, varname, name=F, clu=F)
}
```

##Network (Figure 4E)
```{r}
#network function
net_vis_z <- function(d, i, var){
  colony <- colonyname[i]
  ID <- idlist[i] %>% unlist()
  size <- length(ID)
  caste <- castelist[i] %>% unlist()
  
  d_syn <- d %>% filter(Colony == colony) %>% 
    filter(Sign == "Plus", P < 0.05) %>% select(ID1, ID2, !!sym(var))
  
  id_data <- tibble(ID, caste) %>% 
    mutate(caste_ind = case_when(caste == "BF" ~ 1, caste == "BM" ~ 2, caste == "NB" ~ 3))
  
  set.seed(15)
  pdf(paste0("data/Figure/Figure4/Figure4E/", colony, ".pdf"), width = 3, height = 3)
  net_ss <- graph.data.frame(d = d_syn, vertices = id_data, directed = FALSE)
  E(net_ss)$width <- d_syn %>% pull(!!sym(var)) * 0.1
  col <- c("#AF528D","#54B345","grey")
  V(net_ss)$color <- col[V(net_ss)$caste_ind]
  w <- d_syn %>% pull(!!sym(var))
  c = (1 - ((w-min(w)) / (max(w)-min(w)))) * 0.8
  plot(net_ss, 
       edge.color = rgb(c,c,c), 
       vertex.label = NA, 
       vertex.frame.color = NA,
       vertex.label.color = "black", 
       vertex.label.cex = 0.1, 
       vertex.size = 10)
  dev.off()
}

#run
for(i in 1:5){
  var <- "z_value"
  net_vis_z(zv_syntime, i, var)
}
```


##Network & behavioral cluster (Figure S5D)
```{r}
#network function
net_vis_z_behtype <- function(d, i, var){
  colony <- colonyname[i]
  ID <- idlist[i] %>% unlist()
  size <- length(ID)
  cluster <- d_behtype_mean_d %>% filter(Colony == colony) %>% pull(Cluster)
  
  d_syn <- d %>% filter(Colony == colony) %>% 
    filter(Sign == "Plus", P < 0.05) %>% select(ID1, ID2, !!sym(var))
  
  id_data <- tibble(ID, cluster) %>% mutate(cluster_ind = as.integer(cluster))
  
  set.seed(15)
  pdf(paste0("data/Figure/Figure4/FigureS5D/", colony, ".pdf"), width = 3, height = 3)
  net_ss <- graph.data.frame(d = d_syn, vertices = id_data, directed = FALSE)
  E(net_ss)$width <- d_syn %>% pull(!!sym(var)) * 0.1
  V(net_ss)$color <- colorlist[V(net_ss)$cluster_ind]
  w <- d_syn %>% pull(!!sym(var))
  c = (1 - ((w-min(w)) / (max(w)-min(w)))) * 0.8
  plot(net_ss, edge.color = rgb(c,c,c), 
       vertex.label = NA, vertex.frame.color = NA, vertex.label.color = "black", 
       vertex.label.cex = 0.1, vertex.size = 10)
  dev.off()
}

#run
for(i in 1:5){
  var <- "z_value"
  net_vis_z_behtype(zv_syntime, i, var)
}
```


#3 Statistical analysis using synchrony index

##Add individual attribute
```{r}
zv_syntime_att <- zv_syntime %>% 
  left_join(ID_att_1, by = join_by(Colony, ID1)) %>% 
  left_join(ID_att_2, by = join_by(Colony, ID2)) %>% 
  mutate(Sig = case_when(
    Sign == "Plus" & P < 0.05 ~ "Syn", Sign == "Minus" & P < 0.05 ~ "Asyn", TRUE ~ "NS")
    ) %>% 
  mutate(Sig = factor(Sig, levels = c("Syn", "NS", "Asyn"))) %>% 
  select(Colony, ID1, ID2, Comb, z_value, Sig,
         Caste1, Caste2, Sex1, Sex2, AgeD1, AgeD2, Weight1, Weight2)

zv_syntime_att_t <- zv_syntime_att %>% 
  select(Colony, ID1 = ID2, ID2 = ID1, Comb, z_value, Sign, P, Sig,
         Caste1 = Caste2, Caste2 = Caste1, Sex1 = Sex2, Sex2 = Sex1, 
         AgeD1 = AgeD2, AgeD2 = AgeD1, Weight1 = Weight2, Weight2 = Weight1)
zv_syntime_att_all <- zv_syntime_att %>% bind_rows(zv_syntime_att_t) %>% arrange(Colony, Comb)
```

##Colony-level (Figures S5A & 4F)
```{r}
#significant/ns proportion per colony (figure S5A)
zv_syntime_att_prop <- zv_syntime_att %>% 
  group_by(Colony, Sig) %>% summarise(n = n()) %>% mutate(p = n / sum(n))

ggplot(data = zv_syntime_att_prop, aes(x = Colony, y = p, fill = Sig)) + 
  geom_bar(stat = "identity", position = "stack", width = 0.7) + 
  scale_fill_manual(values = c("#E6855E", "#BEBEBE", "#44A5CB")) + 
  scale_y_continuous(expand = c(0,0)) + 
  theme_classic() + 
  theme(aspect.ratio = 1.5, legend.position = "none",
        axis.title = element_blank(), axis.text = element_blank())
ggsave("data/Figure/Figure4/FigureS5A.pdf", width = 3, height = 4.5, units = "in")

#model
lm_syntime_colony <- lm(data = zv_syntime_att %>% mutate(Colony = as.factor(Colony)), 
                        formula = z_value ~ Colony)
summary(lm_syntime_colony)
#Multiple R-squared:  0.1524,	Adjusted R-squared:  0.1489 

#deviation of EMMs from shuffling control (0)
emm <- emmeans(lm_syntime_colony, ~ Colony)
test_res <- test(emm, null = 0) %>% as_tibble()
write.csv(test_res, row.names = F, "data/statistical_result/Figure4F.csv")

#visualization (figure 4F)
p_lm_syntime_colony <- ggpredict(lm_syntime_colony, terms = c("Colony [all]"))
ggplot() + 
  geom_hline(yintercept = 0, linetype = "dashed", color = "grey") + 
  geom_beeswarm(data = zv_syntime_att, aes(x = Colony, y = z_value, color = Sig), 
                size = 0.8, cex = 1, alpha = 0.75) + 
  geom_errorbar(data = p_lm_syntime_colony, 
                aes(x = x, ymax = predicted, ymin = predicted),
                linewidth = 1, width = 0.5) + 
  scale_color_manual(values = c("#F57446", "#595757", "#669DC8")) + 
  theme_classic() + 
  theme(aspect.ratio = 1.25, legend.position = "none",
        axis.title = element_blank(), axis.text = element_blank())
ggsave("data/Figure/Figure4/Figure4F.pdf", width = 4, height = 5, units = "in")
```

##Caste combination (Figure 4G)
```{r}
#data
zv_syntime_att_castecomb <- zv_syntime_att %>% 
  unite("Caste_comb", c(Caste1, Caste2), sep = "_", remove = F) %>% 
  mutate(Caste_comb = if_else(Caste_comb == "N_B", "B_N", Caste_comb))

#model
lmm_syntime_castecomb <- lmerTest::lmer(
  data = zv_syntime_att_castecomb, formula = z_value ~ Caste_comb + (1|Colony)
)
summary(lmm_syntime_castecomb)
r.squaredGLMM(lmm_syntime_castecomb)
#R2m 0.147708 R2c 0.308726
Anova(lmm_syntime_castecomb, test.statistic = "Chisq")
##x2=213.27, Df=2, p<0.001

#pairwise comparison
emm_res <- emmeans(lmm_syntime_castecomb, pairwise ~ Caste_comb)
pair_res <- pairs(emm_res, adjust = "holm") %>% as_tibble()
write.csv(pair_res, row.names = F, "data/statistical_result/Figure4G.csv")

#visualization
p_lmm_syntime_castecomb <- ggpredict(lmm_syntime_castecomb, terms = c("Caste_comb [all]"))
ggplot() + 
  geom_beeswarm(data = zv_syntime_att_castecomb, aes(x = Caste_comb, y = z_value), 
                size = 0.8, cex = 0.95, alpha = 0.4) + 
  geom_errorbar(data = p_lmm_syntime_castecomb, 
                aes(x = x, ymax = predicted, ymin = predicted),
                size = 1, width = 0.5) + 
  theme_classic() + 
  theme(aspect.ratio = 1.5, legend.position = "none",
        axis.title = element_blank(), axis.text = element_blank())
ggsave("data/Figure/Figure4/Figure4G.pdf", width = 4, height = 6, units = "in")
```

##Sex combination (Figures S5B & S5C)
```{r}
#data construction
zv_syntime_att_sexcomb <- zv_syntime_att %>% 
  unite("Sex_comb", c(Sex1, Sex2), sep = "_", remove = F) %>% 
  mutate(Sex_comb = case_when(
    Sex_comb == "F_BF" ~ "BF_F", 
    Sex_comb == "M_BF" ~ "BF_M", 
    Sex_comb == "F_BM" ~ "BM_F", 
    Sex_comb == "M_BM" ~ "BM_M", 
    Sex_comb == "M_F" ~ "F_M", 
    TRUE ~ Sex_comb
    )) %>% 
  filter(Sex_comb != "BF_BM", Sex_comb != "BM_BF")


#B-NB data
zv_syntime_att_sexcomb_bnb <- zv_syntime_att_sexcomb %>% filter(Caste1 == "B", Caste2 == "N")
lmm_syntime_sexcomb_bnb <- lmerTest::lmer(
  data = zv_syntime_att_sexcomb_bnb, 
  formula = z_value ~ Sex_comb + (1|Colony)
)
r.squaredGLMM(lmm_syntime_sexcomb_bnb)
#R2m 0.021192 R2c 0.2977371
Anova(lmm_syntime_sexcomb_bnb, test.statistic = "Chisq")
##x2=5.5183, Df=3, p=0.1375
emm_res <- emmeans(lmm_syntime_sexcomb_bnb, pairwise ~ Sex_comb)
pair_res <- pairs(emm_res, adjust = "holm") %>% as_tibble()
write.csv(pair_res, row.names = F, "data/statistical_result/FigureS5B.csv")

lmm_syntime_sexcomb_bnb_p <- ggpredict(lmm_syntime_sexcomb_bnb, terms = "Sex_comb")
ggplot() + 
  geom_beeswarm(data = zv_syntime_att_sexcomb_bnb, 
                aes(x = Sex_comb, y = z_value), 
                size = 1.5, cex = 2.5, alpha = 0.4) + 
  geom_errorbar(data = lmm_syntime_sexcomb_bnb_p, 
                aes(x = x, ymax = predicted, ymin = predicted),
                size = 1, width = 0.5) + 
  theme_classic() + 
  theme(aspect.ratio = 1.65, legend.position = "none",
        axis.title = element_blank(), axis.text = element_blank())
ggsave("data/Figure/Figure4/FigureS5B.pdf", width = 2.5, height = 4, units = "in")


#NB-NB data
zv_syntime_att_sexcomb_nbnb <- zv_syntime_att_sexcomb %>% filter(Caste1 == "N", Caste2 == "N")
lmm_syntime_sexcomb_nbnb <- lmerTest::lmer(
  data = zv_syntime_att_sexcomb_nbnb, 
  formula = z_value ~ Sex_comb + (1|Colony)
)
r.squaredGLMM(lmm_syntime_sexcomb_nbnb)
#R2m 0.01183747 R2c 0.2050665
Anova(lmm_syntime_sexcomb_nbnb, test.statistic = "Chisq")
##x2=11.533, Df=2, p=0.003
emm_res <- emmeans(lmm_syntime_sexcomb_nbnb, pairwise ~ Sex_comb)
pair_res <- pairs(emm_res, adjust = "holm") %>% as_tibble()
write.csv(pair_res, row.names = F, "data/statistical_result/FigureS5C.csv")

lmm_syntime_sexcomb_nbnb_p <- ggpredict(lmm_syntime_sexcomb_nbnb, terms = "Sex_comb")
ggplot() + 
  geom_beeswarm(data = zv_syntime_att_sexcomb_nbnb, 
                aes(x = Sex_comb, y = z_value), 
                size = 1, cex = 1.5, alpha = 0.4) + 
  geom_errorbar(data = lmm_syntime_sexcomb_nbnb_p, 
                aes(x = x, ymax = predicted, ymin = predicted),
                size = 1, width = 0.5) + 
  theme_classic() + 
  theme(aspect.ratio = 1.65, legend.position = "none",
        axis.title = element_blank(), axis.text = element_blank())
ggsave("data/Figure/Figure4/FigureS5C.pdf", width = 2.5, height = 4, units = "in")
```

##Behavioral cluster combination (Figure 4H)
```{r}
#data
d_behtype_mean <- read.csv("data/processed_data/representative_cluster.csv")
d_behtype_mean1 <- d_behtype_mean %>% select(Colony, ID1 = ID, Cluster1 = Cluster)
d_behtype_mean2 <- d_behtype_mean %>% select(Colony, ID2 = ID, Cluster2 = Cluster)

zv_syntime_att_behtype_comb <- zv_syntime_att %>% 
  left_join(d_behtype_mean1, by = join_by(Colony, ID1)) %>% 
  left_join(d_behtype_mean2, by = join_by(Colony, ID2)) %>% 
  drop_na(everything()) %>% 
  unite("Cluster_comb", c(Cluster1, Cluster2), sep = "_", remove = F) %>% 
  mutate(Cluster_comb = case_when(
    Cluster_comb == "2_1" ~ "1_2", Cluster_comb == "3_1" ~ "1_3", Cluster_comb == "4_1" ~ "1_4", 
    Cluster_comb == "5_1" ~ "1_5", Cluster_comb == "6_1" ~ "1_6", Cluster_comb == "7_1" ~ "1_7", 
    Cluster_comb == "3_2" ~ "2_3", Cluster_comb == "4_2" ~ "2_4", Cluster_comb == "5_2" ~ "2_5", 
    Cluster_comb == "6_2" ~ "2_6", Cluster_comb == "7_2" ~ "2_7", 
    Cluster_comb == "4_3" ~ "3_4", Cluster_comb == "5_3" ~ "3_5", Cluster_comb == "6_3" ~ "3_6", 
    Cluster_comb == "7_3" ~ "3_7", 
    Cluster_comb == "5_4" ~ "4_5", Cluster_comb == "6_4" ~ "4_6", Cluster_comb == "7_4" ~ "4_7", 
    Cluster_comb == "6_5" ~ "5_6", Cluster_comb == "7_5" ~ "5_7", 
    Cluster_comb == "7_6" ~ "6_7", 
    TRUE ~ Cluster_comb
    )) %>% 
  mutate(Cluster_comb = as.factor(Cluster_comb))

#model
lmm_syntime <- lmerTest::lmer(data = zv_syntime_att_behtype_comb, 
                              formula = z_value ~ 1 + (1|Colony))
syntime_mean <- fixef(lmm_syntime)["(Intercept)"]

lmm_syntime_behtypecomb <- lmerTest::lmer(
  data = zv_syntime_att_behtype_comb, 
  formula = z_value ~ Cluster_comb + (1|Colony)
)
r.squaredGLMM(lmm_syntime_behtypecomb)
#R2m 0.2266085 R2c 0.3715926
Anova(lmm_syntime_behtypecomb, test.statistic = "Chisq")
##x2=348.96, Df=27, p<0.001

#deviation of EMMs from mean (syntime_chance)
emm <- emmeans(lmm_syntime_behtypecomb, ~ Cluster_comb)
test_res <- test(emm, null = syntime_mean) %>% as_tibble()
write.csv(test_res, row.names = F, "data/statistical_result/Figure4H.csv")

#visualization
zv_syntime_att_all_behtype_comb <- zv_syntime_att_all %>% 
  left_join(d_behtype_mean1, by = join_by(Colony, ID1)) %>% 
  left_join(d_behtype_mean2, by = join_by(Colony, ID2)) %>% 
  drop_na(everything()) %>% 
  mutate(Cluster1 = as.factor(Cluster1), Cluster2 = as.factor(Cluster2))
lmm_syntime_all_behtypecomb <- lmerTest::lmer(
  data = zv_syntime_att_all_behtype_comb, 
  formula = z_value ~ Cluster1 * Cluster2 + (1|Colony)
)
lmm_syntime_all_behtypecomb_p <- ggpredict(lmm_syntime_all_behtypecomb, terms = c("Cluster1 [all]", "Cluster2 [all]"))
colorlist <- c("#BE5378", "#CE7754", "#DFC44E", "#54B345", "#3B98B2", "#5F7EB2", "#808080")
ggplot() + 
  geom_hline(yintercept = syntime_mean, linetype = "dashed", color = "grey") + 
  geom_violin(data = zv_syntime_att_all_behtype_comb, 
              aes(x = Cluster1, y = z_value, color = Cluster2),
              position = position_dodge(width = 0.8), width = 0.5, scale = "width") + 
  geom_errorbar(data = lmm_syntime_all_behtypecomb_p, 
                aes(x = x, ymax = predicted, ymin = predicted, color = group), 
                position = position_dodge(width = 0.8), size = 1, width = 0.5) + 
  scale_color_manual(values = colorlist) + 
  theme_classic() + 
  theme(aspect.ratio = 0.25, legend.position = "none",
        axis.title = element_blank(), axis.text = element_blank())
ggsave("data/Figure/Figure4/Figure4H.pdf", width = 16, height = 4, units = "in")
```


#4 Centrality analysis for six 5-day subsets

## 5-day rest/active vector for each id function
```{r}
vector_const_sleact_5 <- function(k){
  
  colony <- colonyname[k]
  id <- idlist[[k]]
  size <- length(id)
  dataset <- dayname[[k]]
  
  #initialize 6 lists per individual
  for(i in 1:size){
    assign(paste0("d_", i, "_list"), vector("list", 6), envir = .GlobalEnv)
  }
  
  #split 6 sets
  for(j in 1:6){
    days_5 <- dataset[((j - 1) * 5 + 1):(j * 5)]
  
    #prepare initial variable for each id
    for(i in 1:size){
      assign(paste0("tmp_", i), "0", envir = .GlobalEnv)
    }
  
    #30-day vector for each id
    for(name in days_5){
      d <- fread(paste0("data/processed_data/locsleep_data/", colony, "/", name, ".csv")) %>% 
        mutate(Time10s = ceiling((Time-0.01)/10)) %>% 
        distinct(Time10s, .keep_all = TRUE) %>% 
        select(-Time, -Time10s)
      
      for(i in 1:size){
        di <- d %>% pull(i) %>% replace_na("0")
        nth <- paste0("tmp_", i)
        assign(nth, c(get(nth), di), envir = .GlobalEnv)
      }
    }
      
    for(i in 1:size){
      tmp <- get(paste0("tmp_", i))[-1]
      di <- tibble(raw = tmp) %>%
        mutate(Sleep = if_else(raw == "S", 1, 0)) %>%
        pull(Sleep)
      d_list_name <- paste0("d_", i, "_list")
      d_list <- get(d_list_name)
      d_list[[j]] <- di
      assign(d_list_name, d_list, envir = .GlobalEnv)
    }
  }
}
```

##Permutation for rest or active function
```{r}
# synchronized times calculation for each 5-day set
syn_time_list <- function(k, ran, ID1, ID2, seg) {
  ID1_seg <- ID1[[seg]]
  ID2_seg <- ID2[[seg]]
  N <- length(ID1_seg)
  if (ran[k] == 1) {
    ID1r <- ID1_seg
  } else {
    ID1r <- ID1_seg[c(ran[k]:N, 1:(ran[k]-1))]
  }
  ID12 <- ID1r + ID2_seg
  Dt <<- sum(ID12 == 2 | ID12 == 0)
}

# permutation (circular) test for each 5-day set
syn_time_pt.test_list <- function(ran, ID1, ID2, nperm, seg) {
  DT <- 0
  for (k in 1:(nperm+1)) {
    syn_time_list(k, ran, ID1, ID2, seg)
    DT <- c(DT, Dt)
  }
  DT <<- DT[-1]
}

# permutation test for all pairs（5 days x 6 sets）
syn_time_pttest_list <- function(colony, nperm){
  size <- idlist[[colony]] %>% length()
  DTs_all <- vector("list", 6)
  set.seed(10)

  for (seg in 1:6){
    pb <- txtProgressBar(min = 1, max = size*(size-1)/2, style = 3)
    DTs <- list()
    N <- length(get(paste0("d_1_list"))[[seg]])
    ran <- sample(2:N, nperm, replace = FALSE)
    ran <- c(1, ran)

    for(i in 1:(size-1)){
      ID1 <- get(paste0("d_", i, "_list"))
      for(j in (i+1):size){
        n <- size*(size-1)/2 - (size-i+1)*(size-i)/2 + (j-i)
        setTxtProgressBar(pb, n)

        ID2 <- get(paste0("d_", j, "_list"))
        syn_time_pt.test_list(ran, ID1, ID2, nperm, seg)
        DTs[[n]] <- DT
      }
    }
    DTs_all[[seg]] <- DTs
  }
  DT_s_list <<- DTs_all
}
```

##Integrate data
```{r}
nperm <- 9999

for (l in 1:5){
  vector_const_sleact_5(l)
  system.time(syn_time_pttest_list(l, nperm))
  size <- idlist[[l]] %>% length()
  n_pair <- size * (size - 1) / 2

  for (seg in 1:6) {
    DT_df <- as.data.frame(do.call(cbind, DT_seg))
    DT_df <- data.frame(DT_seg)
    colnames(DT_df) <- 1:n_pair
    rownames(DT_df) <- paste0("perm_", 1:nperm)
    
    nth <- paste0("DT_", l, "_seg", seg)
    assign(nth, DT_df)
  }
}
#output data: DT_1~5_seg1~6


##Integrate data
trans_tidy_time_seg <- function(obj_name){
  df <- get(obj_name)
  colony   <- as.integer(str_extract(obj_name, "(?<=DT_)\\d+"))
  seg  <- as.integer(str_extract(obj_name, "(?<=_seg)\\d+"))
  df %>% 
    mutate(index = 1:10000) %>% 
    pivot_longer(cols = -index, names_to = "Comb", values_to = "T_val") %>% 
    mutate(Comb = as.integer(Comb)) %>% 
    arrange(Comb) %>% 
    mutate(Colony = colony, Seg = seg) %>% 
    select(Colony, Seg, Comb, index, T_val)
}

df_list <- list()
for (i in 1:5) {
  for (j in 1:6) {
    obj_name <- paste0("DT_", i, "_seg", j)
    df_tidy <- trans_tidy_time_seg(obj_name)
    df_list[[length(df_list) + 1]] <- df_tidy
  }
}
DT_t_t_l <- bind_rows(df_list)
```

##Synchrony index data
```{r}
#p-value summarize function
pv_data_sync5 <- function(d){
  d %>% group_by(Comb) %>% 
    reframe(Value = Syn[index == 1],
            Med = median(Syn),
            Pn = sum(Syn >= Syn[index == 1]) / 10000) %>% 
    mutate(Sign = if_else(Pn < 0.5, "Plus", "Minus")) %>% 
    mutate(P = if_else(Sign == "Plus", Pn * 2, (1 - Pn)*2 + 0.0002)) %>% 
    mutate(logP = -log10(P),
           P_score = if_else(Sign == "Plus", logP, -logP))
}

#run
Colony <- 0; Seg <- 0; Comb <- 0; Value <- 0; Med <- 0; Pn <- 0
Sign <- "0"; P <- 0; logP <- 0; P_score <- 0
pv0 <- tibble(Colony, Seg, Comb, Value, Med, Pn, Sign, P, logP, P_score)

for(i in 1:5){
  for(j in 1:6){
    d <- DT_t_t_l %>% filter(Colony == i, Seg == j) %>% mutate(Syn = T_val)
    pv <- pv_data_sync5(d) %>% mutate(Colony = i, Seg = j)
    pv0 <- pv0 %>% bind_rows(pv)
  }
}
pv_sync5 <- pv0 %>% slice(-1)

#synchrony index data (z_value = synchrony index)
DT_t_t_l_sim_s <- DT_t_t_l %>% 
  group_by(Colony, Seg, Comb) %>% 
  filter(index != 1) %>% 
  summarise(Mean = mean(T_val), SD = sd(T_val))
zv_sync5 <- pv_sync5 %>% 
  left_join(DT_t_t_l_sim_s, join_by(Colony, Seg, Comb)) %>% 
  mutate(z_value = (Value - Mean) / SD) %>% 
  left_join(ID_att_comb, join_by(Colony, Comb))
```

##Strength data
```{r}
strength_sync <- list()
combs <- zv_sync5 %>% distinct(Colony, Seg)

for (k in 1:nrow(combs)) {
  col <- combs$Colony[k]
  seg <- combs$Seg[k]

  df_sub <- zv_sync5 %>% filter(Colony == col, Seg == seg)

  g <- graph_from_data_frame(
    d = df_sub %>% select(ID1, ID2, z_value),
    directed = FALSE
  )

  strength_vals <- strength(g, weights = E(g)$z_value)
  
  result <- tibble(
    Colony = col,
    Seg = seg,
    ID = names(strength_vals),
    Strength = strength_vals
  )

  strength_sync[[length(strength_sync) + 1]] <- result
}

d_behtype_mean <- read.csv("data/processed_data/representative_cluster.csv") %>% 
  mutate(Cluster = as.factor(Cluster))
d_behtype_mean %>% group_by(Cluster) %>% summarise(n = n()) #16 14 11 13 7 31 9

strength_sync <- bind_rows(strength_sync) %>% 
  left_join(ID_att %>% select(ID, Caste, Sex), join_by(ID)) %>% 
  left_join(d_behtype_mean %>% select(ID, Cluster), join_by(ID)) %>% 
  mutate(
    ID = factor(ID, levels = id),
    Caste = case_when(Sex == "BF" ~ "BF", Sex == "BM" ~ "BM", TRUE ~ Caste)
  ) %>% 
  arrange(Colony, Seg, ID) %>% 
  group_by(Colony, Seg) %>% 
  mutate(Seg = factor(Seg, levels = rev(1:6)), 
         yj.Strength = yeojohnson(Strength)$x.t) %>% #transform
  ungroup() %>% 
  unite(col = "ColSeg", Colony, Seg, sep = "_", remove = F) #colony x segment column

#mean data
strength_sync_mean <- strength_sync %>% 
  group_by(Colony, ID) %>% 
  summarise(mean_yj = mean(yj.Strength),
            sd_yj = sd(yj.Strength)) %>% 
  arrange(Colony, ID)
```

##Individual differences in strength (Figure S5E)
```{r}
#plot

#zscore heatmap
for(i in 1:5){
  colony <- colonyname[i]
  size <- idlist[[i]] %>% length()
  p <- ggplot(strength_sync %>% filter(Colony == i), 
              aes(x = ID, y = Seg, fill = yj.Strength)) + 
    geom_tile() + 
    scale_fill_gradientn(colors = colorRampPalette(RdYlBu)(n=100)) + 
    theme_void() + 
    theme(aspect.ratio = 10 / size, legend.position = "none")
  ggsave(plot = p, 
         paste0("data/Figure/Figure4/FigureS5E_heatmap_", colony, ".pdf"), 
         width = 4 * size / 10, height = 4, units = "in")
}

#zscore mean heatmap
for(i in 1:5){
  colony <- colonyname[i]
  size <- idlist[[i]] %>% length()
  p <- ggplot(strength_sync_mean %>% filter(Colony == i), 
              aes(x = ID, y = Seg, fill = mean_yj)) + 
    geom_tile() + 
    scale_fill_gradientn(colors = colorRampPalette(RdYlBu)(n=100)) + 
    theme_void() + 
    theme(aspect.ratio = 10 / size / 6, legend.position = "none")
  ggsave(plot = p, 
         paste0("data/Figure/Figure4/FigureS5E_mean_heatmap_", colony, ".pdf"), 
         width = 4 * size / 10, height = 4 / 6, units = "in")
}

#zscore meansd plot
for(i in 1:5){
  colony <- colonyname[i]
  p <- ggplot(strength_sync %>% filter(Colony == i), 
              aes(x = ID, y = yj.Strength, color = Caste)) + 
    geom_hline(yintercept = 0, linetype = "dashed", color = "grey") + 
    geom_jitter(size = 1, width = 0.25, height = 0, alpha = 0.5) + 
    stat_summary(fun = mean, geom = "crossbar", width = 0.5) + 
    stat_summary(fun.data = "mean_sdl", geom = "errorbar", width = 0) + 
    scale_color_manual(values = c("#AF528D","#54B345","dimgrey")) + 
    labs(x = "ID", y = "Strength") + 
    theme_classic() + 
    theme(aspect.ratio = 0.5, legend.position = "none")
  ggsave(plot = p, 
         paste0("data/Figure/Figure4/FigureS5E_meansdplot_", colony, ".pdf"), 
         width = 8, height = 4, units = "in")
}


#Bayesian LMM
set.seed(100)
blmm_syn_str_id <- brm(
  formula = Strength ~ 0 + ColSeg + (1 | ID),
  data = strength_sync,
  family = gaussian(),
  chains = 4, iter = 4000, warmup = 1000
)
waic(blmm_syn_str_id) #4240.6

#Individual estimate + CI
re_blmm_syn_str_id <- posterior_samples(blmm_syn_str_id, pars = "^r_ID\\[")

re_blmm_syn_str_id_summary <- re_blmm_syn_str_id %>%
  pivot_longer(cols = everything(), names_to = "param", values_to = "value") %>%
  mutate(ID = sub("r_ID\\[([^,]+),Intercept\\]", "\\1", param),
         ID = factor(ID, levels = id)) %>%
  group_by(ID) %>%
  summarise(
    Estimate = mean(value),
    Lower95 = quantile(value, 0.025),
    Upper95 = quantile(value, 0.975),
    Lower99 = quantile(value, 0.005),
    Upper99 = quantile(value, 0.995),
    Lower999 = quantile(value, 0.0005),
    Upper999 = quantile(value, 0.9995)
  ) %>%
  mutate(
    Sig = case_when(Lower999 > 0 ~ "H***", Upper999 < 0 ~ "L***",
                    Lower99 > 0 ~ "H**", Upper99 < 0 ~ "L**",
                    Lower95 > 0 ~ "H*", Upper95 < 0 ~ "L*",
                    TRUE ~ "ns")
  )
write.csv(re_blmm_syn_str_id_summary, row.names = F, 
          "data/statistical_result/FigureS5E_test.csv")
```

##Cluster differences in strength (Figure S5F)
```{r}
#zscore plot
ggplot(strength_sync %>% filter(is.na(Cluster) == FALSE), 
       aes(x = Cluster, y = yj.Strength, color = Cluster)) + 
  geom_hline(yintercept = 0, linetype = "dashed", color = "grey") + 
  geom_beeswarm(cex = 1.5, size = 1.2, alpha = 0.5) + 
  stat_summary(fun = mean, geom = "crossbar", width = 0.5, fatten = 2) + 
  scale_color_manual(values = colorlist) + 
  theme_classic() + 
  theme(aspect.ratio = 1, legend.position = "none",
        axis.title = element_blank(), axis.text = element_blank())
ggsave("data/Figure/Figure4/FigureS5F.pdf", 
       width =  4, height = 4, units = "in")


#LMM
lmm_syn_str_clu <- lmerTest::lmer(
  data = strength_sync,
  Strength ~ Cluster + (1|ID) + (1|ColSeg)
)
r2(lmm_syn_str_clu) #Conditional R2: 0.691 Marginal R2: 0.117

#EMM
emm <- emmeans(lmm_syn_str_clu, ~ Cluster)
grand_mean <- mean(summary(emm)$emmean)
test_res <- test(emm, null = grand_mean) %>% as_tibble()
write.csv(test_res, row.names = F, 
          "data/statistical_result/FigureS5F_test.csv")
```




