---
title: "Figure 6"
output: html_notebook
---

#1 Permutation function

##Move event data function
```{r}
move_event_data <- function(k){
  
  colony <- colonyname[k]
  dayvec <- dayname[[k]]
  id <- idlist[[k]]
  size <- length(id)
  
  #ID, Time, From_To data for 30-day
  ID <- "0"; Time <- 0; From_To <- "0"
  d_f <- tibble(ID, Time, From_To)
  
  for(i in 1:30){
    name <- dayvec[i]
    
    d_f0 <- fread(paste0("data/processed_data/beh_event_data/", colony, "/", name, ".csv")) %>% 
      #remove disturbance phase
      filter(Disturb == 0) %>% 
      #extract move events of clear from_to
      group_by(ID) %>% 
      mutate(Loc_lag = lag(Loc, n=1)) %>% 
      drop_na(Loc_lag) %>% filter(Loc != "", Loc_lag != "") %>% 
      ungroup() %>% 
      select(ID, Time, From = Loc_lag, To = Loc) %>% 
      unite("From_To", c(From, To), sep = "_", remove = F) %>% 
      #transform time column
      mutate(Day = i,
             Time = Time + 86400 * (i-1),
             ID = factor(ID, levels = id)) %>% 
      select(ID, Time, From_To) %>% 
      arrange(Time, ID)
    
    d_f <- d_f %>% bind_rows(d_f0)
  }
  d_f %>% slice(-1) %>% mutate(ID = factor(ID, levels = id))
}
```

##Permutation for follow behavior function
```{r}
follow_pttest_t <- function(d, colony, nperm){
  
  id <- idlist[colony] %>% unlist()
  size <- length(id)
  
  #set random number
  set.seed(10)
  
  #result list
  DF_i <- list(); DF_j <- list()
  DF_ijs <- list(); DF_jis <- list()
  
  #dyad data
  id_pairs <- expand.grid(i = 1:(size - 1), j = (2:size)) %>% filter(i < j)
  
  #parallel setting
  cl <- makeCluster(detectCores() - 1)
  clusterExport(cl, list("d_f", "d", "id", "size", "mutate", "lag", "inner_join", "%>%", "runif", "set.seed"))
  clusterEvalQ(cl, {library(dplyr)})

  #parallel per dyad
  results <- parLapply(cl, 1:nrow(id_pairs), function(idx) { 
    i <- id_pairs$i[idx]
    j <- id_pairs$j[idx]
    
    res_i <- list(); res_j <- list()
    res_ij <- list(); res_ji <- list()
    
    #id i&j data
    d_ij <- d %>% filter(ID == id[i] | ID == id[j]) %>% arrange(Time)
    d_i <- d_ij %>% filter(ID == id[i])
    d_i_time <- d_i %>% pull(Time)
    d_i_fromto <- d_i %>% pull(From_To)
    d_j <- d_ij %>% filter(ID == id[j])
    
    #time window for 3-sec interval & index including one or two ID
    d_ij_ind <- d_ij %>% 
      mutate(Time_lag = lag(Time, n=1),
             dif_Time = Time - Time_lag,
             dif = if_else(dif_Time > 3, 1, 0),
             dif = if_else(row_number() == 1, 1, dif),
             Index = as.integer(cumsum(dif))) %>% 
      group_by(Index) %>% 
      mutate(n_dist = n_distinct(ID)) %>% 
      ungroup()
    Index <- d_ij_ind %>% pull(Index)
    n_dist <- d_ij_ind %>% pull(n_dist)
    
    #shuffle indices
    shuffle_indices <- sample(1:(length(d_i_fromto)-1), nperm-1)
    d_i_fromto_list <- list(0)
    
    # npermの繰り返し
    Df_i <- numeric(nperm); Df_j <- numeric(nperm)
    Df_ij <- numeric(nperm); Df_ji <- numeric(nperm)
    
    for(k in 1:nperm){
      if(k == 1){
        d_ij_t <- d_ij
      } else {
        d_i_fromto_list[[k-1]] <- d_i_fromto[(seq_along(d_i_fromto) + shuffle_indices[k-1]) %% length(d_i_fromto) + 1]
        d_i_t <- tibble(ID = id[i], Time = d_i_time, From_To = d_i_fromto_list[[k-1]])
        d_ij_t <- bind_rows(d_i_t, d_j) %>% arrange(Time)
      }
      
      #remove time window with only one individual 
      d_ij_ind <- d_ij_t %>% 
        mutate(Index = Index, n_dist = n_dist) %>% 
        filter(n_dist > 1) %>% select(-n_dist)
      
      suppressWarnings({  #ignore warning about numerous numbers
        d_ij_ind_sm <- d_ij_ind %>% 
          #extract dyad with same Index & same From_To
          group_by(Index) %>% 
          inner_join(d_ij_ind, by = c("Index", "From_To"), suffix = c("_1", "_2")) %>% 
          filter(ID_1 != ID_2, ID_1 == id[i]) %>% 
          #identify directional follow times
          mutate(dif_Time = Time_1 - Time_2) %>% 
          filter(dif_Time >= -3, dif_Time <= 3, dif_Time != 0) %>% 
          mutate(n_ij = if_else(dif_Time < 0, 1, 0),
                 n_ji = if_else(dif_Time > 0, 1, 0))
        
        #save sum of result
        Df_i[k] <- id[i]; Df_j[k] <- id[j]
        Df_ij[k] <- sum(d_ij_ind_sm$n_ij)
        Df_ji[k] <- sum(d_ij_ind_sm$n_ji)
      })
      
      res_i <- list(Df_i); res_j <- list(Df_j)
      res_ij <- list(Df_ij)
      res_ji <- list(Df_ji)
    }
    
    return(list(res_i, res_j, res_ij, res_ji))
  })
  
  #result
  DF_i <<- unlist(lapply(results, `[[`, 1))
  DF_j <<- unlist(lapply(results, `[[`, 2))
  DF_ijs <<- unlist(lapply(results, `[[`, 3))
  DF_jis <<- unlist(lapply(results, `[[`, 4))
  
  #end parallel action
  stopCluster(cl)
  #close connection
  closeAllConnections()
}
```

##Run function
```{r}
nperm <- 1000

for(l in 1:5){
  d_f <- move_event_data(l)
  system.time(follow_pttest_t(d_f, 1, nperm))
  
  size <- idlist[l] %>% unlist() %>% length()
  DFt <- tibble(DF_i, DF_j, DF_ijs, DF_jis)
  nth <- paste0("DFt_", l)
  assign(nth, DFt)
}
#output data: DFt_1~5
```

##Integrate data
```{r}
trans_tidy_follow <- function(d, i){
  colony <- colonyname[i]
  id <- idlist[i] %>% unlist()
  d %>% 
    select(ID_i = DF_i, ID_j = DF_j, Nij = DF_ijs, Nji = DF_jis) %>% 
    mutate(ID_i = factor(ID_i, levels = id),
           ID_j = factor(ID_j, levels = id)) %>% 
    arrange(ID_i, ID_j) %>% 
    group_by(ID_i, ID_j) %>% 
    mutate(Index = 1:1000) %>% 
    ungroup() %>% 
    pivot_longer(cols = c(-ID_i, -ID_j, -Index), names_to = "Direction", values_to = "N") %>% 
    mutate(ID_1 = ID_i, ID_2 = ID_j) %>% 
    mutate(ID_i = if_else(Direction == "Nji", ID_2, ID_i),
           ID_j = if_else(Direction == "Nji", ID_1, ID_j)) %>% 
    unite("Perm", c(ID_i, ID_j), sep = "_", remove = F) %>% 
    arrange(ID_i, ID_j, Index) %>% 
    mutate(Colony = colony) %>% 
    select(Colony, Perm, ID_i, ID_j, Index, N)
}

DFt_1_l <- trans_tidy_follow(DFt_1, 1)
DFt_2_l <- trans_tidy_follow(DFt_2, 2)
DFt_3_l <- trans_tidy_follow(DFt_3, 3)
DFt_4_l <- trans_tidy_follow(DFt_4, 4)
DFt_5_l <- trans_tidy_follow(DFt_5, 5)
DFt_t_l <- bind_rows(DFt_1_l, DFt_2_l, DFt_3_l, DFt_4_l, DFt_5_l)
```

##Dyad information
```{r}
ID_att <- read.csv("data/pre_info_data/individual_information.csv") %>% mutate(Birth = as.Date(Birth))

ID_att_1 <- ID_att %>% select(Colony, ID1 = ID, Caste1 = Caste, Sex1 = Sex, AgeD1 = AgeD, Weight1 = Weight)
ID_att_2 <- ID_att %>% select(Colony, ID2 = ID, Caste2 = Caste, Sex2 = Sex, AgeD2 = AgeD, Weight2 = Weight)
```

#2 Visualization of p-value & follow index

##p-value data
```{r}
#p-value summarize function
pv_data_fol <- function(d, i){
  colony <- colonyname[i]
  d_i <- d %>% filter(Colony == colony)
  perm <- d_i %>% distinct(Perm, .keep_all = F) %>% pull(Perm)
  d_i %>% 
    group_by(ID_i, ID_j, Perm) %>% 
    summarise(Value = N[Index == 1],
              Med = median(N),
              Pn = sum(N >= N[Index == 1]) / 1000) %>% 
    mutate(Sign = if_else(Pn < 0.5, "Plus", "Minus")) %>% 
    mutate(P = if_else(Sign == "Plus", Pn * 2, (1 - Pn)*2 + 0.002)) %>% 
    mutate(logP = -log10(P),
           P_score = if_else(Sign == "Plus", logP, -logP)) %>% 
    mutate(Colony = colony) %>% 
    mutate(Perm = factor(Perm, levels = perm)) %>% 
    select(Colony, Perm, ID_i, ID_j, 
           Value, Med, Pn, Sign, P, logP, P_score) %>% 
    arrange(Perm)
}

#run
Colony <- "0"; Perm <- "0"; ID_i <- "0"; ID_j <- "0"
Value <- 0; Med <- 0; Pn <- 0; Sign <- "0"; P <- 0; logP <- 0; P_score <- 0
pv0 <- tibble(Colony, Perm, ID_i, ID_j, Value, Med, Pn, Sign, P, logP, P_score)
for(i in 1:5){
  pv <- pv_data_fol(DFt_t_l, i)
  pv0 <- pv0 %>% bind_rows(pv)
}
pv_folt <- pv0 %>% slice(-1)
pv_folt
```

##Follow index data
```{r}
#example of frequency distribution in simulated value (Figure 6B)
DFt_t_l_LOC1_OC1 <- DFt_t_l %>% filter(Colony == "LOC", ID_i == "LOC1", ID_j == "OC1")
DFt_t_l_LOC1_OC1 %>% filter(Index != 1) %>% 
  group_by(Perm) %>% summarise(mean = mean(N), sd = sd(N))
#mean 186.4014	sd 27.33093
DFt_t_l_LOC1_OC1 %>% filter(Index == 1) %>% pull(N)
#220
ggplot(data = DFt_t_l_LOC1_OC1 %>% filter(Index != 1)) + 
  geom_histogram(aes(x = N)) + 
  scale_y_continuous(expand = c(0,0)) + 
  theme_classic() + 
  theme(aspect.ratio = 1, axis.title = element_blank(), axis.text = element_blank())
ggsave("data/Figure/Figure6/Figure6B.pdf", width = 3, height = 3, units = "in")

#follow index data (yj.z_value = follow index)
DFt_t_l_sim_s <- DFt_t_l %>% 
  group_by(Colony, Perm) %>% 
  filter(Index != 1) %>% 
  summarise(Mean = mean(N), SD = sd(N))
zv_folt <- pv_folt %>% 
  left_join(DFt_t_l_sim_s, join_by(Colony, Perm)) %>% 
  mutate(z_value = (Value - Mean) / SD) %>% 
  #Yeo-Johnson transformation & standardization (all)
  mutate(yj.z_value = bestNormalize::yeojohnson(z_value)$x.t)

#calculate shuffling control value
zv_vect <- pv_folt %>% 
  left_join(DFt_t_l_sim_s, join_by(Colony, Perm)) %>% 
  mutate(z_value = (Value - Mean) / SD) %>% 
  pull(z_value)
zv_vect_yj <- bestNormalize::yeojohnson(zv_vect)
zv_zero <- (0 - zv_vect_yj$mean) / zv_vect_yj$sd
```

##Heatmap for follow index (Figure 6C)
```{r}
#matrix function
zv_mat_data_fol <- function(d, i){
  colony <- colonyname[i]
  id <- idlist[i] %>% unlist()
  size <- length(id)
  d0 <- tibble(ID1 = id, ID2 = id, x = rep(NA, size))
  d %>% 
    filter(Colony == colony) %>% 
    select(ID1 = ID_i, ID2 = ID_j, x = yj.z_value) %>% 
    bind_rows(d0) %>% 
    mutate(ID1 = as.factor(ID1), ID2 = as.factor(ID2)) %>% 
    arrange(ID1, ID2) %>% 
    pivot_wider(names_from = ID2, values_from = x) %>% 
    column_to_rownames(var = "ID1")
}

#pheatmap function
zv_hm_fol <- function(d, i, name, clu){
  colony <- colonyname[i]
  ID <- idlist[i] %>% unlist()
  Caste <- castelist[i] %>% unlist()
  anno <- tibble(ID, Caste) %>% column_to_rownames(var = "ID")
  Caste_col <- c("#AF528D","#54B345","grey")
  names(Caste_col) <- c("BF","BM","NB")
  anno_color <- list(Caste = Caste_col)
  RdYlBu <- c("#313695", "#4575b4", "#74add1", "#abd9e9", "#e0f3f8", "#ffffbf", 
              "#fee090", "#fdae61", "#f46d43", "#d73027","#a50026")
  
  val <- d %>% rowSums(1:length(ID)) %>% as_tibble() %>% pull(value)
  callback = function(hc, mat){
    dend = reorder(as.dendrogram(hc), wts = -val)
    as.hclust(dend)
  }
  
  pheatmap(
    mat = d,
    na_col = "white",
    annotation_row = anno, 
    annotation_col = anno,
    annotation_colors = anno_color,
    cellheight = 10, cellwidth = 10,
    border_color = "NA",
    breaks = seq(zv_min, zv_max, length.out = 100),
    color = colorRampPalette(RdYlBu)(n=100),
    annotation_legend = FALSE,
    annotation_names_row = name, annotation_names_col = name,
    show_rownames = name, show_colnames = name,
    legend = name,
    cluster_rows = clu, cluster_cols = clu,
    clustering_method = "ward.D2",
    clustering_callback = callback,
    filename = paste0("data/Figure/Figure6/Figure6C/", colony, "_name", name, "_clu", clu, ".pdf")
    )
}

#perform
for(i in 1:5){
  colony <- colonyname[i]
  d <- zv_folt %>% filter(Colony == colony)
  zv_mat <- zv_mat_data_fol(d, i)
  zv_hm_fol(zv_mat, i, varname, name=F, clu=T)
  zv_hm_fol(zv_mat, i, varname, name=T, clu=T)
}
```

##Heatmap for p-value (Figure 6D)
```{r}
#id sequence from figure 6C
id_1_follow_r <- c("2B33", "2B42", "M14", "2B32", "2B35", "M4", "2B46", "2B51", "2B44", "2B38", "2B43",
                "2B50", "2B39", "2B48", "2B34", "2B36", "2B47", "2B45", "2B40", "2B41", "2B37", "2B49")
id_1_follow_c <- c("M4", "M14", "2B33", "2B32", "2B35", "2B34", "2B39", "2B45", "2B42", "2B38", "2B41",
                "2B37", "2B49", "2B51", "2B47", "2B48", "2B40", "2B50", "2B36", "2B43", "2B46", "2B44")
id_2_follow_r <- c("2D31", "2D29", "2D34", "2D26", "2D13", "2D12", "2D24", "2D35", "2D30",
                   "2D32", "2D33", "2D11", "H11", "2D19", "2D16", "2D28", "2D23")
id_2_follow_c <- c("2D13", "2D34", "2D32", "2D33", "2D26", "2D24", "2D29", "2D28", "2D31",
                   "2D16", "2D19", "2D23", "2D11", "2D35", "2D12", "2D30", "H11")
id_3_follow_r <- c("HG24", "HG10", "HG23", "HG5", "HG19", "HG22", "HG14", "HG21", "HG16", "HG4",
                   "HG12", "HG7", "HG11", "HG13", "HG15", "HG8", "HG18", "G16", "H13")
id_3_follow_c <- c("HG24", "HG14", "HG21", "HG23", "HG10", "HG19", "HG16", "HG11", "HG18", "HG22",
                   "HG5", "HG15", "HG13", "HG12", "HG7", "HG8", "HG4", "G16", "H13")
id_4_follow_r <- c("HT14", "HT11", "HT5", "HT4", "HT12", "HT9", "HT13", "HT7", "HT10", "HT1", "HT6",
                  "HT2", "HT3", "HT17", "HT15", "HT19", "HT21", "HT16", "L63", "R21", "HT20", "HT18")
id_4_follow_c <- c("HT3", "HT16", "HT21", "HT15", "HT17", "HT20", "HT18", "HT19", "HT5", "R21", "L63",
                  "HT11", "HT12", "HT10", "HT7", "HT13", "HT9", "HT14", "HT4", "HT2", "HT6", "HT1")
id_5_follow_r <- c("LOC16", "LOC14", "LOC15", "LOC10", "LOC7", "LOC5", "LOC13", "LOC9", 
                   "LOC8", "LOC11", "LOC4", "LOC17", "OC1", "LOC2", "LOC6", "LOC22",
                   "LOC23", "LOC1", "LOC20", "LOC18", "LOC19", "L55")
id_5_follow_c <- c("LOC7", "LOC14", "LOC13", "LOC8", "LOC9", "LOC11", "LOC15", "LOC10",
                   "LOC16", "LOC17", "LOC18", "LOC4", "LOC1", "LOC5", "LOC6", "LOC2",
                   "LOC20", "LOC23", "LOC22", "LOC19", "L55", "OC1")
idlist_follow_r <- list(id_1_follow_r, id_2_follow_r, id_3_follow_r, id_4_follow_r, id_5_follow_r)
idlist_follow_c <- list(id_1_follow_c, id_2_follow_c, id_3_follow_c, id_4_follow_c, id_5_follow_c)


#matrix function
pv_mat_data_fol <- function(d, i, var, order_r, order_c){
  colony <- colonyname[i]
  id <- idlist[i] %>% unlist()
  size <- length(id)
  d0 <- tibble(ID1 = id, ID2 = id, x = rep(NA, size))
  d %>% 
    filter(Colony == colony) %>% 
    select(ID1 = ID_i, ID2 = ID_j, x = !!sym(var)) %>% 
    bind_rows(d0) %>% 
    mutate(ID1 = factor(ID1, levels = order_r),
           ID2 = factor(ID2, levels = order_c)) %>% 
    arrange(ID1, ID2) %>% 
    pivot_wider(names_from = ID2, values_from = x) %>% 
    column_to_rownames(var = "ID1")
}

#pheatmap function
pv_hm_fol <- function(d, i, name, clu){
  colony <- colonyname[i]
  ID <- idlist[i] %>% unlist()
  Caste <- castelist[i] %>% unlist()
  anno <- tibble(ID, Caste) %>% column_to_rownames(var = "ID")
  Caste_col <- c("#AF528D","#54B345","grey")
  names(Caste_col) <- c("BF","BM","NB")
  anno_color <- list(Caste = Caste_col)
  RdYlBu <- c("#313695", "#4575b4", "#74add1", "#abd9e9", "#e0f3f8", "#ffffbf", 
              "#fee090", "#fdae61", "#f46d43", "#d73027","#a50026")
  pheatmap(
    mat = d,
    na_col = "white",
    annotation_row = anno, 
    annotation_col = anno,
    annotation_colors = anno_color,
    cellheight = 10, cellwidth = 10,
    border_color = "NA",
    breaks = seq(log10(0.002), -log10(0.002), length.out=100),
    color = colorRampPalette(RdYlBu)(n=100),
    annotation_legend = FALSE,
    annotation_names_row = name, annotation_names_col = name,
    show_rownames = name, show_colnames = name,
    legend = name,
    cluster_rows = clu, cluster_cols = clu,
    filename = paste0("data/Figure/Figure6/Figure6D/", colony, "_name", name, "_clu", clu, ".pdf")
    )
}

#run
for(i in 1:5){
  d <- pv_folt
  var <- "P_score"
  order_r <- idlist_follow_r[i] %>% unlist()
  order_c <- idlist_follow_c[i] %>% unlist()
  pv_mat <- pv_mat_data_fol(d, i, var, order_r, order_c)
  pv_hm_fol(pv_mat, i, varname, name=F, clu=F)
}
```

##Network (Figure 6E)
```{r}
#network function
net_vis_z_folt <- function(d, i, var){
  colony <- colonyname[i]
  ID <- idlist[i] %>% unlist()
  size <- length(ID)
  caste <- castelist[i] %>% unlist()
  
  d_sig <- d %>% filter(Colony == colony) %>% 
    filter(Sign == "Plus", P < 0.05) %>% select(ID_j, ID_i, !!sym(var))
  
  id_data <- tibble(ID, caste) %>% 
    mutate(caste_ind = case_when(caste == "BF" ~ 1, caste == "BM" ~ 2, caste == "NB" ~ 3))
  
  set.seed(15)
  pdf(paste0("data/Figure/Figure6/Figure6E/", colony, ".pdf"), width = 3, height = 3)
  net_ss <- graph_from_data_frame(d = d_sig, vertices = id_data, directed = TRUE)
  w <- d_sig %>% pull(!!sym(var)) %>% rescale(c(0.01,1))
  wt <- (1 - ((w-min(w)) / (max(w)-min(w))))
  E(net_ss)$width <- w
  col <- c("#AF528D","#54B345","grey")
  V(net_ss)$color <- col[V(net_ss)$caste_ind]
  c = wt * 0.8
  plot(net_ss, 
       edge.color = rgb(c,c,c), edge.curved = 0.3, edge.arrow.size = 0.1, 
       vertex.label = NA, vertex.frame.color = NA, vertex.label.color = "black", 
       vertex.label.cex = 0.1, vertex.size = 10)
  dev.off()
}

#run
for(i in 1:5){
  var <- "z.yj.z_value"
  net_vis_z_folt(zv_folt, i, var)
}
```

##Network & behavioral cluster (Figure S6F)
```{r}
#network function
net_vis_z_behtype_fol <- function(d, i, var, varname){
  colony <- colonyname[i]
  ID <- idlist[i] %>% unlist()
  size <- length(ID)
  cluster <- d_behtype_mean_d %>% filter(Colony == colony) %>% pull(Cluster)
  
  d_sig <- d %>% filter(Colony == colony) %>% 
    filter(Sign == "Plus", P < 0.05) %>% select(ID_j, ID_i, !!sym(var))
  
  id_data <- tibble(ID, cluster) %>% mutate(cluster_ind = as.integer(cluster))
  
  set.seed(15)
  pdf(paste0("data/Figure/Figure6/FigureS6F/", colony, ".pdf"), width = 3, height = 3)
  net_ss <- graph_from_data_frame(d = d_sig, vertices = id_data, directed = TRUE)
  w <- d_sig %>% pull(!!sym(var)) %>% rescale(c(0.01,1))
  wt <- (1 - ((w-min(w)) / (max(w)-min(w))))
  E(net_ss)$width <- w
  V(net_ss)$color <- colorlist[V(net_ss)$cluster_ind]
  c = wt * 0.8
  plot(net_ss, 
       edge.color = rgb(c,c,c), edge.curved = 0.3, edge.arrow.size = 0.1, 
       vertex.label = NA, vertex.frame.color = NA, vertex.label.color = "black", 
       vertex.label.cex = 0.1, vertex.size = 10)
  dev.off()
}

#run
for(i in 1:5){
  var <- "yj.z_value"
  net_vis_z_behtype_fol(zv_folt, i, var)
}
```


#3 Statistical analysis using follow index

##Add individual attribute
```{r}
#data
zv_folt_att <- zv_folt %>% 
  select(Colony, ID1 = ID_i, ID2 = ID_j, yj.z_value, Sign, P) %>% 
  left_join(ID_att_1, by = join_by(Colony, ID1)) %>% 
  left_join(ID_att_2, by = join_by(Colony, ID2)) %>% 
  mutate(Sig = case_when(
    Sign == "Plus" & P < 0.05 ~ "Posive", Sign == "Minus" & P < 0.05 ~ "Negative", TRUE ~ "NS")
    ) %>% 
  mutate(Sig = factor(Sig, levels = c("Posive", "NS", "Negative"))) %>% 
  select(Colony, ID1, ID2, yj.z_value, Sig,
         Caste1, Caste2, Sex1, Sex2, AgeD1, AgeD2, Weight1, Weight2)

#add relative age & relative body weight data for nonbreeders
ID_att_nb <- ID_att %>% 
  filter(Caste == "N") %>% 
  group_by(Colony) %>% 
  mutate(log_Age = log(AgeD), zAge = scale(log_Age), zWeight = scale(Weight)) %>% 
  ungroup() %>% 
  select(Colony, ID, zAge, zWeight)
ID_att_nb1 <- ID_att_nb %>% select(Colony, ID1 = ID, zAge1 = zAge, zWeight1 = zWeight)
ID_att_nb2 <- ID_att_nb %>% select(Colony, ID2 = ID, zAge2 = zAge, zWeight2 = zWeight)

zv_folt_att <- zv_folt_att %>% 
  left_join(ID_att_nb1, join_by(Colony, ID1)) %>% 
  left_join(ID_att_nb2, join_by(Colony, ID2))
```

##Colony-level (Figures S6A)
```{r}
#significant/ns proportion per colony (Figure S6A)
zv_folt_att_prop <- zv_folt_att %>% 
  group_by(Colony, Sig) %>% summarise(n = n()) %>% mutate(p = n / sum(n))

ggplot(data = zv_folt_att_prop, aes(x = Colony, y = p, fill = Sig)) + 
  geom_bar(stat = "identity", position = "stack", width = 0.7) + 
  scale_fill_manual(values = c("#E6855E", "#BEBEBE", "#44A5CB")) + 
  scale_y_continuous(expand = c(0,0)) + 
  theme_classic() + 
  theme(aspect.ratio = 1.5, legend.position = "none",
        axis.title = element_blank(), axis.text = element_blank())
#1 sp 0.1450, ns 0.7208, sn 0.1342
#2 sp 0.3713, ns 0.4963, sn 0.1324
#3 sp 0.4883, ns 0.4562, sn 0.0556
#4 sp 0.4134, ns 0.5411, sn 0.0454
#5 sp 0.2143, ns 0.7186, sn 0.0671
ggsave("data/Figure/Figure6/FigureS6A.pdf", width = 3, height = 4.5, units = "in")
```


##Caste difference (Figure 6F)
```{r}
#data
zv_folt_att_castecomb <- zv_folt_att %>% 
  select(Colony, ID1, ID2, yj.z_value, Caste1, Caste2) %>% 
  unite("CasteComb", c(Caste2, Caste1), sep = "_", remove = F)

#model
lmm_fft_castecomb <- lmerTest::lmer(
  data = zv_folt_att_castecomb %>% mutate(CasteComb = as.factor(CasteComb)), 
  formula = yj.z_value ~ CasteComb + (1|Colony)
)
r.squaredGLMM(lmm_fft_castecomb)
#R2m 0.09014773 R2c 0.1838073
Anova(lmm_fft_castecomb, test.statistic = "Chisq")
#Sum Sq   Df  Pr(>F)
#CasteComb  220.79  3  < 2.2e-16 ***

#deviation of EMMs from shuffling control (zv_zero)
emm <- emmeans(lmm_fft_castecomb, ~ CasteComb)
test_res <- test(emm, null = zv_zero) %>% as_tibble()
write.csv(test_res, row.names = F, "data/statistical_result/Figure6F_test.csv")

#multiple comparisons of EMMs
emm_pair <- emmeans(lmm_fft_castecomb, pairwise ~ CasteComb)
pairs_res <- pairs(emm_pair, adjust = "holm")
write.csv(pairs_res, row.names = F, "data/statistical_result/Figure6F_pairs.csv")

#plot
p_lmm_fft_castecomb_zero <- ggpredict(lmm_fft_castecomb_zero, terms = c("CasteComb [all]"))
ggplot() + 
  geom_hline(yintercept = zv_zero, linetype = "dashed", color = "grey") + 
  geom_beeswarm(data = zv_folt_att_castecomb, 
                aes(x = CasteComb, y = yj.z_value), 
                size = 0.4, cex = 0.55, alpha = 0.4) + 
  geom_errorbar(data = p_lmm_fft_castecomb_zero, 
                aes(x = x, ymax = predicted + zv_zero, ymin = predicted + zv_zero),
                size = 1, width = 0.5) + 
  theme_classic() + 
  theme(aspect.ratio = 1.5, legend.position = "none",
        axis.title = element_blank(), axis.text = element_blank())
ggsave("data/Figure/Figure6/Figure6F.pdf", width = 4, height = 6, units = "in")
```

##BF vs BM in B-B data (Figure S6B)
```{r}
#data
zv_folt_att_qk <- zv_folt_att %>% filter(Caste1 == "B", Caste2 == "B")

#paired t-test
bb_bf_vect_t <- zv_folt_att_qk %>% filter(Sex2 == "BF") %>% pull(yj.z_value) %>% as.vector()
bb_bm_vect_t <- zv_folt_att_qk %>% filter(Sex2 == "BM") %>% pull(yj.z_value) %>% as.vector()
shapiro.test(bb_bf_vect_t - bb_bm_vect_t)
t.test(bb_bf_vect_t, bb_bm_vect_t, paired = T)
#mean dif=-1.073586, t=-2.3048, df=4, p=0.08251

#plot
ggplot(zv_folt_att_qk, aes(x = Sex2, y = yj.z_value, group = Colony)) + 
  geom_point(size = 1) + 
  geom_line() + 
  labs(x = "Follower", y = "Follow index") + 
  theme_classic() + 
  theme(aspect.ratio = 2.1,
        axis.title = element_blank(), axis.text = element_blank())
ggsave("data/Figure/Figure6/FigureS6B.pdf", width = 2, height = 4, units = "in")
```

##Sex combination

###B-NB data (Figure S6C)
```{r}
#data
zv_folt_att_nbb <- zv_folt_att %>% 
  filter(Caste1 == "N", Caste2 == "B") %>% 
  unite("SexComb", c(Sex2, Sex1), sep = "_", remove = F)

#model
lmm_fft_nbb_sexcomb_all <- lmerTest::lmer(
  data = zv_folt_att_nbb, 
  formula = yj.z_value ~ SexComb + (1|Colony)
)
r.squaredGLMM(lmm_fft_nbb_sexcomb_all)
#R2m 0.003220203 R2c 0.09806367
Anova(lmm_fft_nbb_sexcomb_all, test.statistic = "Chisq")
##x2=0.6475, Df=3, p=0.8855

#multiple comparisons of EMMs
emm_pair <- emmeans(lmm_fft_nbb_sexcomb_all, pairwise ~ SexComb)
pairs_res <- pairs(emm_pair, adjust = "holm") %>% as_tibble()
write.csv(pairs_res, row.names = F, "data/statistical_result/FigureS6C.csv")

#plot
lmm_fft_nbb_sexcomb_all_p <- ggpredict(lmm_fft_nbb_sexcomb_all, terms = "SexComb")
ggplot() + 
  geom_beeswarm(data = zv_folt_att_nbb, 
                aes(x = SexComb, y = yj.z_value), 
                size = 1.5, cex = 2.5, alpha = 0.4) + 
  geom_errorbar(data = lmm_fft_nbb_sexcomb_all_p, 
                aes(x = x, ymax = predicted, ymin = predicted),
                size = 1, width = 0.5) + 
  theme_classic() + 
  theme(aspect.ratio = 1.65, legend.position = "none",
        axis.title = element_blank(), axis.text = element_blank())
ggsave("data/Figure/Figure6/FigureS6C.pdf", width = 2.5, height = 4, units = "in")
```

###NB-B data (Figure S6D)
```{r}
#data
zv_folt_att_bnb <- zv_folt_att %>% 
  filter(Caste1 == "B", Caste2 == "N") %>% 
  unite("SexComb", c(Sex2, Sex1), sep = "_", remove = F)

#model
lmm_fft_bnb_sexcomb_all <- lmerTest::lmer(
  data = zv_folt_att_bnb, 
  formula = yj.z_value ~ SexComb + (1|Colony)
)
r.squaredGLMM(lmm_fft_bnb_sexcomb_all)
#R2m 0.03622761 R2c 0.1207381
Anova(lmm_fft_bnb_sexcomb_all, test.statistic = "Chisq")
##x2=7.5319, Df=3, p=0.05674

#multiple comparisons of EMMs
emm_pair <- emmeans(lmm_fft_bnb_sexcomb_all, pairwise ~ SexComb)
pairs_res <- pairs(emm_pair, adjust = "holm") %>% as_tibble()
write.csv(pairs_res, row.names = F, "data/statistical_result/FigureS6D.csv")

#plot
lmm_fft_bnb_sexcomb_all_p <- ggpredict(lmm_fft_bnb_sexcomb_all, terms = "SexComb")
ggplot() + 
  geom_beeswarm(data = zv_folt_att_bnb, 
                aes(x = SexComb, y = yj.z_value), 
                size = 1.5, cex = 2.5, alpha = 0.4) + 
  geom_errorbar(data = lmm_fft_bnb_sexcomb_all_p, 
                aes(x = x, ymax = predicted, ymin = predicted),
                size = 1, width = 0.5) + 
  theme_classic() + 
  theme(aspect.ratio = 1.65, legend.position = "none",
        axis.title = element_blank(), axis.text = element_blank())
ggsave("data/Figure/Figure6/FigureS6D.pdf", width = 2.5, height = 4, units = "in")
```

###NB-NB data (Figure S6E)
```{r}
#data
zv_folt_att_nbnb <- zv_folt_att %>% 
  filter(Caste1 == "N", Caste2 == "N") %>% 
  unite("SexComb", c(Sex2, Sex1), sep = "_", remove = F)

#model
lmm_fft_nbnb_sexcomb_all <- lmerTest::lmer(
  data = zv_folt_att_nbnb, 
  formula = yj.z_value ~ SexComb + (1|Colony)
)
r.squaredGLMM(lmm_fft_nbnb_sexcomb_all)
#R2m 0.03608966 R2c 0.168823
Anova(lmm_fft_nbnb_sexcomb_all, test.statistic = "Chisq")
##x2=69.269, Df=3, p<0.001

#multiple comparisons of EMMs
emm_pair <- emmeans(lmm_fft_nbnb_sexcomb_all, pairwise ~ SexComb)
pairs_res <- pairs(emm_pair, adjust = "holm") %>% as_tibble()
write.csv(pairs_res, row.names = F, "data/statistical_result/FigureS6E.csv")

#plot
lmm_fft_nbnb_sexcomb_all_p <- ggpredict(lmm_fft_nbnb_sexcomb_all, terms = "SexComb")
lmm_fft_nbnb_sexcomb_all_p
ggplot() + 
  geom_beeswarm(data = zv_folt_att_nbnb, 
                aes(x = SexComb, y = yj.z_value), 
                size = 0.4, cex = 0.9, alpha = 0.4) + 
  geom_errorbar(data = lmm_fft_nbnb_sexcomb_all_p, 
                aes(x = x, ymax = predicted, ymin = predicted),
                size = 1, width = 0.5) + 
  theme_classic() + 
  theme(aspect.ratio = 1.65, legend.position = "none",
        axis.title = element_blank(), axis.text = element_blank())
ggsave("data/Figure/Figure6/FigureS6E.pdf", width = 2.5, height = 4, units = "in")
```

##BF vs BM related to nonbreeder's weight in B-NB data (Figure 6G)
```{r}
#model
lmm_fft_nbb_zwei <- lmerTest::lmer(
  data = zv_folt_att_nbb, 
  formula = yj.z_value ~ poly(zWeight1,2) * Sex2 + (1|Colony)
)
r.squaredGLMM(lmm_fft_nbb_zwei)
#R2m 0.1064013 R2c 0.2047355
model_res <- summary(lmm_fft_nbb_zwei)$coefficients %>% as.data.frame()
write.csv(model_res, row.names = T, "data/statistical_result/Figure6G.csv")

Anova(lmm_fft_nbb_zwei, test = "Chisq")
#Chisq Df Pr(>Chisq)
#poly(zWeight1, 2)      16.1073  2  0.0003179 ***
#Sex2                    0.1446  1  0.7037155    
#poly(zWeight1, 2):Sex2  8.2319  2  0.0163104 *

#plot
lmm_fft_nbb_zwei_p <- ggpredict(lmm_fft_nbb_zwei, terms = c("zWeight1 [all]", "Sex2"))
ggplot() + 
  geom_point(data = zv_folt_att_nbb, 
             aes(x = zWeight1, y = yj.z_value, color = Sex2, shape = Sex2)) + 
  geom_line(data = lmm_fft_nbb_zwei_p, aes(x = x, y = predicted, color = group)) + 
  geom_ribbon(data = lmm_fft_nbb_zwei_p, 
              aes(x = x, ymin = conf.low, ymax = conf.high, fill = group), alpha = 0.1) + 
  scale_color_manual(values = c("#AF528D", "#54B345")) + 
  scale_fill_manual(values = c("#AF528D", "#54B345")) + 
  theme_classic() + 
  theme(aspect.ratio = 1, legend.position = "none",
        axis.title = element_blank(), axis.text = element_blank())
ggsave("data/Figure/Figure6/Figure6G.pdf", 
       width = 4, height = 4, units = "in")
```

##BF vs BM related to nonbreeder's age in B-NB data (Figure S6F)
```{r}
#model
lmm_fft_nbb_zage <- lmerTest::lmer(
  data = zv_folt_att_nbb, 
  formula = yj.z_value ~ poly(zAge1,2) * Sex2 + (1|Colony)
)
r.squaredGLMM(lmm_fft_nbb_zage)
#R2m 0.1254181 R2c 0.2241243
model_res <- summary(lmm_fft_nbb_zage)$coefficients %>% as.data.frame()
write.csv(model_res, row.names = T, "data/statistical_result/FigureS6F.csv")

Anova(lmm_fft_nbb_zage, test = "Chisq")
#Chisq Df Pr(>Chisq)
#poly(zAge1, 2)      24.8486  2   4.02e-06 ***
#Sex2                 0.1483  1     0.7001    
#poly(zAge1, 2):Sex2  4.5834  2     0.1011

#plot
lmm_fft_nbb_zage_p <- ggpredict(lmm_fft_nbb_zage, terms = c("zAge1 [all]", "Sex2"))
ggplot() + 
  geom_point(data = zv_folt_att_nbb, 
             aes(x = zAge1, y = yj.z_value, color = Sex2, shape = Sex2)) + 
  geom_line(data = lmm_fft_nbb_zage_p, aes(x = x, y = predicted, color = group)) + 
  geom_ribbon(data = lmm_fft_nbb_zage_p, 
              aes(x = x, ymin = conf.low, ymax = conf.high, fill = group), alpha = 0.1) + 
  scale_color_manual(values = c("#AF528D", "#54B345")) + 
  scale_fill_manual(values = c("#AF528D", "#54B345")) + 
  theme_classic() + 
  theme(aspect.ratio = 1, legend.position = "none",
        axis.title = element_blank(), axis.text = element_blank())
ggsave("data/Figure/Figure6/FigureS6F.pdf", width = 4, height = 4, units = "in")
```

##Behavioral cluster combination (Figure 6H)
```{r}
#data
d_behtype_mean <- read.csv("data/processed_data/representative_cluster.csv")
d_behtype_mean1 <- d_behtype_mean %>% select(Colony, ID1 = ID, Cluster1 = Cluster)
d_behtype_mean2 <- d_behtype_mean %>% select(Colony, ID2 = ID, Cluster2 = Cluster)

zv_folt_att_behtypeperm <- zv_folt_att %>% 
  left_join(d_behtype_mean1, by = join_by(Colony, ID1)) %>% 
  left_join(d_behtype_mean2, by = join_by(Colony, ID2)) %>% 
  drop_na(Cluster1, Cluster2) %>% 
  unite("ClusterComb", c(Cluster2, Cluster1), sep = "_", remove = F) %>% 
  mutate(ClusterComb = as.factor(ClusterComb),
         Cluster1 = as.factor(Cluster1),
         Cluster2 = as.factor(Cluster2))

#model
lmm_fft_behtypecomb <- lmerTest::lmer(
  data = zv_folt_att_behtypecomb, 
  formula = yj.z_value ~ ClusterComb + (1|Colony)
)
r.squaredGLMM(lmm_fft_behtypecomb)
#R2m 0.156148 R2c 0.2475557
Anova(lmm_fft_behtypecomb, test.statistic = "Chisq")
#Chisq Df Pr(>Chisq)
#Cluster_perm  405.73 48  < 2.2e-16 ***

#deviation of EMMs from mean (0)
emm <- emmeans(lmm_fft_behtypecomb, ~ ClusterComb)
test_res <- test(emm, null = 0) %>% as_tibble()
write.csv(test_res, row.names = F, "data/statistical_result/Figure6H.csv")

#plot
lmm_fft_behtypecomb_all <- lmerTest::lmer(
  data = zv_folt_att_behtypecomb, 
  formula = yj.z_value ~ Cluster1 * Cluster2 + (1|Colony)
)
lmm_fft_behtypecomb_all_p <- ggpredict(lmm_fft_behtypecomb_all, terms = c("Cluster2 [all]", "Cluster1 [all]"))
colorlist <- c("#BE5378", "#CE7754", "#DFC44E", "#54B345", "#3B98B2", "#5F7EB2", "#808080")
ggplot() + 
  geom_hline(yintercept = 0, linetype = "dashed", color = "grey") + 
  geom_violin(data = zv_folt_att_behtypecomb, 
              aes(x = Cluster2, y = yj.z_value, color = Cluster1),
              position = position_dodge(width = 0.8), width = 0.5, scale = "width") + 
  geom_errorbar(data = lmm_fft_behtypecomb_all_p, 
                aes(x = x, ymax = predicted, ymin = predicted, color = group), 
                position = position_dodge(width = 0.8), size = 1, width = 0.5) + 
  scale_color_manual(values = colorlist) + 
  theme_classic() + 
  theme(aspect.ratio = 0.25, legend.position = "none",
        axis.title = element_blank(), axis.text = element_blank())
ggsave("data/Figure/Figure6/Figure6H.pdf", width = 16, height = 4, units = "in")
```


