---
title: "Figure 2-3"
output: html_notebook
---

#Figure 2

##Data formatting
```{r}
#unify data
id <- c(id_2b, id_2d, id_hg, id_ht, id_loc)
Colony <- "0"; Day <- "0"; ID <- "0"
Sum_Garbage <- 0; Sum_Nest <- 0; Sum_Others <- 0; Sum_Sleep <- 0; Sum_Toilet <- 0
N_Garbage <- 0; N_Nest <- 0; N_Others <- 0; N_Sleep <- 0; N_Toilet <- 0
Mean_Garbage <- 0; Mean_Nest <- 0; Mean_Others <- 0; Mean_Sleep <- 0; Mean_Toilet <- 0
SD_Garbage <- 0; SD_Nest <- 0; SD_Others <- 0; SD_Sleep <- 0; SD_Toilet <- 0
CV_Garbage <- 0; CV_Nest <- 0; CV_Others <- 0; CV_Sleep <- 0; CV_Toilet <- 0 
dt <- tibble(Colony, Day, ID, 
            Sum_Garbage, Sum_Nest, Sum_Others, Sum_Sleep, Sum_Toilet, 
            N_Garbage, N_Nest, N_Others, N_Sleep, N_Toilet, 
            Mean_Garbage, Mean_Nest, Mean_Others, Mean_Sleep, Mean_Toilet, 
            SD_Garbage, SD_Nest, SD_Others, SD_Sleep, SD_Toilet, 
            CV_Garbage, CV_Nest, CV_Others, CV_Sleep, CV_Toilet)

for(i in colonyname){
  dataset <- dayname[[i]]
  for(name in dataset){
    d <- read.csv(paste0("data/processed_data/beh_event_summary_data/", i, "/", name, ".csv")) %>% 
      mutate(Colony = i, Day = name)
    dt <- bind_rows(dt, d)
  }
}

dt <- dt %>% slice(-1) %>% 
  mutate(across(c(-c(1:3,19:28)), ~ replace_na(.x, 0))) %>%  
  mutate(Day = factor(Day, levels = dayname %>% unlist()),
         ID = factor(ID, levels = id)) %>% 
  arrange(Day, ID) %>% 
  mutate(Day = as.character(Day), ID = as.character(ID))

#relationship between n_other and movement tendency
dt_move <- dt %>% 
  select(N_Sleep, N_Nest, N_Toilet, N_Garbage, N_Others) %>% 
  mutate(N_Total = rowSums(.)) %>% 
  mutate(N_Others_Total = N_Others / N_Total)
cor.test(dt_move$N_Others, dt_move$N_Total)
##r=0.9911168, t = 412.11, df = 3058, p-value < 2.2e-16
dt_move %>% summarize(N_Others_Total_mean = mean(N_Others_Total), N_Others_Total_sd = sd(N_Others_Total))
##mean=0.7997396	sd=0.06448304	

#add variables(relative to non-rest stay event)
dta <- dt %>% 
  select(-c(19:28)) %>% #remove SD column and CV column
  mutate(Sum_Active = Sum_Garbage + Sum_Nest + Sum_Others + Sum_Toilet) %>% 
  mutate(
    Sum_Garbage_relA = Sum_Garbage / Sum_Active,
    Sum_Nest_relA = Sum_Nest / Sum_Active,
    Sum_Others_relA = Sum_Others / Sum_Active,
    Sum_Toilet_relA = Sum_Toilet / Sum_Active,
    N_Garbage_relA = N_Garbage / Sum_Active,
    N_Nest_relA = N_Nest / Sum_Active,
    N_Others_relA = N_Others / Sum_Active,
    N_Toilet_relA = N_Toilet / Sum_Active
    )

#confirm existence of 0
dt %>% filter(Sum_Garbage == 0)
dt %>% filter(Sum_Nest == 0)
dt %>% filter(Sum_Others == 0)
dt %>% filter(Sum_Sleep == 0)
dt %>% filter(Sum_Toilet == 0)
dta
#data transformation
dta_rel <- dta %>%
  select(-19) %>% #remove active column
  #add minimum value to columns including 0 (Sum,N,Mean,Sum_relA,N_relA_Garbage)
  mutate(across(c(4,9,14,19,23), ~ . + min(.[.!=0]), .names = "f.{.col}")) %>% 
  #Box-Cox transformation per colony
  group_by(Colony) %>% 
  mutate(across(c(26,4:7,27,9:12,28,14:17,29,19:21,30,23:25), 
                ~ bcPower(., powerTransform(.)$lambda), .names = "bc.{.col}")) %>% 
  ungroup() %>%
  #standardization (within day and colony)
  select(1:3,32:54) %>% 
  group_by(Colony, Day) %>% 
  mutate(across(c(2:24), ~ c(scale(.)), .names = "z.{.col}")) %>% 
  ungroup() %>% select(1:3,27:49)

#behavioral parameter data (23 var)
colnames(dta_rel) <- c(
  "Colony","Day","ID",
  "Sum_Garbage","Sum_Nest","Sum_Others","Sum_Sleep","Sum_Toilet",
  "N_Garbage","N_Nest","N_Others","N_Sleep","N_Toilet",
  "Mean_Garbage","Mean_Nest","Mean_Others","Mean_Sleep","Mean_Toilet",
  "Sum_relA_Garbage","Sum_relA_Nest","Sum_relA_Others","Sum_relA_Toilet",
  "N_relA_Garbage","N_relA_Nest","N_relA_Others","N_relA_Toilet"
)
dta_rel <- dta_rel %>% select(1:3,7,12,17,5,20,10,24,15,6,21,11,25,16,8,22,13,26,18,4,19,9,23,14)
```


##Clustering

###UMAP (first) (Figure S2A)
```{r}
config <- umap.defaults
config$random_state <- 20
config$n_neighbors <- 5

d_um_t0 <- umap(dta_rel[,4:26], config = config)
d_um_t0_t <- tibble(as.data.frame(d_um_t0$layout))
ggplot(d_um_t0_t) + 
  geom_point(aes(x=V1, y=V2), size=0.2) + 
  theme_void() + theme(aspect.ratio = 1)
ggsave("data/Figure/Figure2/FigureS2A.pdf", width = 5, height = 5, units = "in")

dta_rel_umap <- dta_rel %>% bind_cols(d_um_t0_t)
dta_rel_umap %>% filter(ID == "HT7")
ggplot() + 
  geom_point(data = d_um_t0_t , aes(x=V1, y=V2), size=0.2) + 
  geom_point(data = d_um_t0_t %>% filter(ID == "HT7") , aes(x=V1, y=V2), size=0.2, color="red") + 
  theme_void() + theme(aspect.ratio = 1)
#Outlier = HT7 (only one individual)

#remove HT7 from data
dta_rel2 <- dta_rel %>% filter(ID != "HT7")

id_ht0 <- id_ht[-9]
caste_ht0 <- c("BF", "BM", rep("NB", 19))
```

###UMAP (second)
```{r}
d_um_t1 <- umap(dta_rel2[,4:26], config = config)
d_um_t1_t <- tibble(as.data.frame(d_um_t1$layout))
ggplot(d_um_t1_t) + 
  geom_point(aes(x=V1, y=V2), size=0.2) + 
  theme_void() + theme(aspect.ratio = 1)
```

###UMAP density map
```{r}
spectral <- c("#5E4FA2","#3288BD","#66C2A5","#ABDDA4","#E6F598","#FFFFBF",
              "#FEE08B","#FDAE61","#F46D43","#D53E4F","#9E0142")

#density map (color) visualization function
umap_densmap_color <- function(d, bw){
  da_um_t_ggdens <<- ggplot(d, aes(x=V1, y=V2)) + 
    stat_density_2d(aes(fill=after_stat(density)), contour = F, geom = "raster", 
                    h = c(bw, bw), n = 500) + 
    scale_fill_gradientn(colors = spectral) + 
    theme_void() + 
    theme(aspect.ratio = 1, legend.position = "none") +
    scale_x_continuous(expand = c(0,0)) +
    scale_y_continuous(expand = c(0,0))
}

#density map (grey) visualization function
umap_densmap_grey <- function(d, bw){
  da_um_t_ggdens_grey <<- ggplot(d, aes(x=V1, y=V2)) + 
    stat_density_2d(aes(fill=after_stat(density)), contour = F, geom = "raster", 
                    h = c(bw, bw), n = 500) + 
    scale_fill_distiller(palette = "Greys") + 
    theme_void() + 
    theme(aspect.ratio = 1, legend.position = "none") +
    scale_x_continuous(expand = c(0,0)) +
    scale_y_continuous(expand = c(0,0))
}

#visualize density map
se=20; ne=5; bw=2.8
umap_densmap_color(d_um_t1_t, bw)
umap_densmap_grey(d_um_t1_t, bw)
ggsave(da_um_t_ggdens_grey, 
       "data/Figure/Figure2/Figure2C.png", width=500, height=500, units="px")

#fix image data
cimg.limit.openmp()
im_p <- load.image("data/Figure/Figure2/Figure2C.png")
im_p <- rm.alpha(im_p)
im_pg <- grayscale(im_p)
save.image(im_pg, "data/Figure/Figure2/Figure2C_fix.png")
```


###Watershed segmentation (Figures 2B-2C)

watershed segmentation in matlab
matlab code

A = imread("data/Figure/Figure2/Figure2C_fix.png");
B = imcomplement(A);
C = watershed(B);
imwrite(C, "data/Figure/Figure2/Figure2C_fix_ws.png")


```{r}
#confirm result watershed segmentation
im_p_ws <- load.image("data/Figure/Figure2/Figure2C_fix_ws.png")
plot(as.raster(im_p_ws))
im_p_ws2 <- im_p_ws
im_p_ws2[im_p_ws2 > 0] = 1 #only boundary lines
save.image(im_p_ws2, "data/Figure/Figure2/Figure2C_fix_ws2.png")
im_p_ws3 <- image_read("data/Figure/Figure2/Figure2C_fix_ws2.png")
im_p_ws3 <- image_transparent(im_p_ws3, 'white') #transparent processing
image_write(im_p_ws3, "data/Figure/Figure2/Figure2C_fix_ws_tp.png")
image_write(im_p_ws3, "data/Figure/Figure2/Figure2C.pdf")
im_p_ws_d <- data.frame(im_p_ws[,,1,1]) #digitize watershed result

#transform watershed result
im_p_ws_d_cl <- im_p_ws_d %>% 
  mutate(X = 1:500) %>% 
  pivot_longer(cols = -X, names_to = "Yn", values_to = "Value") %>% 
  mutate(Y = rep(500:1, 500)) %>% 
  select(1,4,3)
cluster <- im_p_ws_d_cl %>% group_by(Value) %>% group_indices() 
im_p_ws_d_cl <- im_p_ws_d_cl %>% mutate(Cluster = cluster) %>% select(-Value)
ggplot(im_p_ws_d_cl) + geom_tile(aes(x=X, y=Y, fill=Cluster)) + theme_classic() + theme(aspect.ratio = 1)

#confirm illustration range
xlim <- ggplot_build(da_um_t_ggdens)$layout$panel_params[[1]]$x.range
ylim <- ggplot_build(da_um_t_ggdens)$layout$panel_params[[1]]$y.range
xmin <- xlim[1]; xmax <- xlim[2]
ymin <- ylim[1]; ymax <- ylim[2]

#pixelate UMAP data
d_um_t_500 <- d_um_t1_t %>% 
  mutate(X500 = (V1 - xmin) * 500 / (xmax - xmin),
         Y500 = (V2 - ymin) * 500 / (ymax - ymin)) %>% 
  mutate(X = as.integer(ceiling(X500)), 
         Y = as.integer(ceiling(Y500)))

#add cluster column UMAP data referring to watershed result
d_um_t_500_cl <- d_um_t_500 %>% left_join(im_p_ws_d_cl)
ggplot(d_um_t_500_cl) + 
  geom_point(aes(x=V1, y=V2, color=as.character(Cluster))) + 
  theme_void() + 
  theme(aspect.ratio = 1, legend.position = "none")

#modify boundary sample in UMAP data (change the same cluster as the nearest sample)
x <- d_um_t_500_cl$X500; y <- d_um_t_500_cl$Y500
Clus <- d_um_t_500_cl$Cluster
for(i in 1:length(x)){
  if(Clus[i] != 1){next}
  if(Clus[i] == 1){
    d_um_t_500_cl_dis <- d_um_t_500_cl %>% 
      mutate(dis = (X-x[i])^2 + (Y-y[i])^2)
    Clus[i] <- d_um_t_500_cl_dis %>% 
      filter(Cluster != 1) %>% 
      slice(which.min(dis)) %>% 
      pull(Cluster)
  }
}
d_um_t_500_cl <- d_um_t_500_cl %>% mutate(Cluster = Clus)
#3->1, 2->2, 4->3, 8->4, 6->5, 5->6, 7->7

#modify cluster name
d_um_t_500_cl <- d_um_t_500_cl %>% 
  mutate(Cluster = case_when(
    Cluster == 3 ~ "1", Cluster == 2 ~ "2", Cluster == 4 ~ "3", Cluster == 8 ~ "4", 
    Cluster == 6 ~ "5", Cluster == 5 ~ "6", Cluster == 7 ~ "7")) %>% 
  mutate(Cluster = factor(Cluster, levels = c("1","2","3","4","5","6","7")))

#visualize the clustering result
colorlist <- c("#BE5378","#CE7754","#DFC44E","#54B345","#3B98B2","#5F7EB2","#808080")
ggplot(d_um_t_500_cl) + 
  geom_point(aes(x=V1, y=V2, color=Cluster), size=1) + 
  scale_color_manual(values = colorlist) + 
  theme_void() + 
  theme(aspect.ratio = 1, legend.position = "none")
ggsave("data/Figure/Figure2/Figure2B.pdf", width = 5, height = 5, units = "in")
```

###Combine cluster and UMAP data into behavioral parameter data
```{r}
dta_rel_umap <- bind_cols(dta_rel2, d_um_t_500_cl %>% select(Cluster, V1, V2))
dta_rel_umap$Cluster <- as.character(dta_rel_umap$Cluster)

dta_rel_umap_layout <- dta_rel_umap %>% select(Colony, Day, ID, V1, V2)
write.csv(dta_rel_umap_layout, "data/processed_data/umap_layout.csv", row.names = F)
```


##Relationships between clusters and behavioral parameters

###Visualization for relationship between parameters and UMAP data (Figures 2D-2G, S3)
```{r}
umap_var_spectral <- function(d, varlist){
  for(var in varlist){
    x <- d %>% pull(var)
    ggplot(d) + 
      geom_point(aes(x=V1, y=V2, color=x), size=1) + 
      scale_color_gradientn(colors = spectral) + 
      theme_void() + 
      theme(aspect.ratio = 1,
            legend.position = "none")
    ggsave(paste0("data/Figure/Figure2/FigureS3/", var, ".pdf"), width = 4, height = 4, units = "in")
  }
}

d <- dta_rel_umap
varlist <- c(
  c("Sum_Sleep", "N_Sleep", "Mean_Sleep"),
  c("Sum_Nest", "Sum_relA_Nest", "N_Nest", "N_relA_Nest", "Mean_Nest"),
  c("Sum_Others", "Sum_relA_Others", "N_Others", "N_relA_Others", "Mean_Others"),
  c("Sum_Toilet", "Sum_relA_Toilet", "N_Toilet", "N_relA_Toilet", "Mean_Toilet"),
  c("Sum_Garbage", "Sum_relA_Garbage", "N_Garbage", "N_relA_Garbage", "Mean_Garbage")
  )
umap_var_spectral(d, varlist)
```

###Comparison of each parameter across clusters (Figures 2D-G)
```{r}
#data
dta_rel_umap_para <- dta_rel_umap %>% mutate(Cluster = as.factor(Cluster))

#model N (Rest)
lm_para_N_Sleep <- lm(data = dta_rel_umap_para, formula = N_Sleep ~ Cluster)
summary(lm_para_N_Sleep)
#Multiple R-squared:  0.5425,	Adjusted R-squared:  0.5416
Anova(lm_para_N_Sleep, test.statistic = "Chisq")
##x2=1555.9, Df=6, p<0.001
emm_res <- emmeans(lm_para_N_Sleep, pairwise ~ Cluster)
pair_res <- pairs(emm_res, adjust = "holm") %>% as_tibble()
write.csv(pair_res, row.names = F, "data/statistical_result/Figure2D.csv")

#model Mean (Rest)
lm_para_Mean_Sleep <- lm(data = dta_rel_umap_para, formula = Mean_Sleep ~ Cluster)
summary(lm_para_Mean_Sleep)
#Multiple R-squared:  0.4011,	Adjusted R-squared:  0.3999
Anova(lm_para_Mean_Sleep, test.statistic = "Chisq")
##x2=1157.8, Df=6, p<0.001
emm_res <- emmeans(lm_para_Mean_Sleep, pairwise ~ Cluster)
pair_res <- pairs(emm_res, adjust = "holm") %>% as_tibble()
write.csv(pair_res, row.names = F, "data/statistical_result/Figure2E.csv")

#model Sum (Rest)
lm_para_Sum_Sleep <- lm(data = dta_rel_umap_para, formula = Sum_Sleep ~ Cluster)
summary(lm_para_Sum_Sleep)
#Multiple R-squared:  0.4056,	Adjusted R-squared:  0.4045
Anova(lm_para_Sum_Sleep, test.statistic = "Chisq")
##x2=1173.7, Df=6, p<0.001
emm_res <- emmeans(lm_para_Sum_Sleep, pairwise ~ Cluster)
pair_res <- pairs(emm_res, adjust = "holm") %>% as_tibble()
write.csv(pair_res, row.names = F, "data/statistical_result/Figure2F.csv")

#model N relative (Other)
lm_para_N_relA_Others <- lm(data = dta_rel_umap_para, formula = N_relA_Others ~ Cluster)
summary(lm_para_N_relA_Others)
#Multiple R-squared:  0.4316,	Adjusted R-squared:  0.4305
Anova(lm_para_N_relA_Others, test.statistic = "Chisq")
##x2=1184.8, Df=6, p<0.001
emm_res <- emmeans(lm_para_N_relA_Others, pairwise ~ Cluster)
pair_res <- pairs(emm_res, adjust = "holm") %>% as_tibble()
write.csv(pair_res, row.names = F, "data/statistical_result/Figure2G.csv")
```

###Visualization for relationship between parameters and clusters (Figure S3)
```{r}
#data
dta_rel_umap_mean <- dta_rel_umap %>% 
  select(-c(V1,V2)) %>% 
  group_by(Cluster) %>% 
  summarise(across(4:26, mean))
varlevel <- c(
  c("Sum_Sleep", "N_Sleep", "Mean_Sleep"), 
  c("Sum_Nest", "N_Nest", "Mean_Nest", "Sum_relA_Nest", "N_relA_Nest"), 
  c("Sum_Others", "N_Others", "Mean_Others", "Sum_relA_Others", "N_relA_Others"), 
  c("Sum_Toilet", "N_Toilet", "Mean_Toilet", "Sum_relA_Toilet", "N_relA_Toilet"), 
  c("Sum_Garbage", "N_Garbage", "Mean_Garbage", "Sum_relA_Garbage", "N_relA_Garbage")
)
dta_rel_umap_mean_l <- dta_rel_umap_mean %>% 
  pivot_longer(cols = -1, names_to = "Variable", values_to = "Mean") %>% 
  mutate(Variable = factor(Variable, levels = varlevel))

dta_rel_umap_sd <- dta_rel_umap %>% 
  select(-c(V1,V2)) %>% 
  group_by(Cluster) %>% 
  summarise(across(4:26, sd, .names = "{.col}"))
dta_rel_umap_sd_l <- dta_rel_umap_sd %>% 
  pivot_longer(cols = -1, names_to = "Variable", values_to = "SD") %>% 
  mutate(Variable = factor(Variable, levels = varlevel))

dta_rel_umap_clu_n <- dta_rel_umap %>% group_by(Cluster) %>% summarise(n = n())

dta_rel_umap_meansd_l <- dta_rel_umap_mean_l %>% 
  bind_cols(dta_rel_umap_sd_l %>% select(SD)) %>% 
  left_join(dta_rel_umap_clu_n, by = join_by(Cluster)) %>% 
  select(Cluster, n, Variable, Mean, SD)

#plot
d <- dta_rel_umap_meansd_l %>% filter(grepl(pattern = "Sleep", x = Variable))
ggplot(d, aes(x = Variable, y = Mean, color = Cluster)) + 
  geom_errorbar(aes(ymax = Mean + SD, ymin = Mean - SD), 
                position = position_dodge(width=0.8), width = 0.5) + 
  geom_point(position = position_dodge(width=0.8), size = 3) + 
  scale_color_manual(values = colorlist) + 
  scale_x_discrete(expand = c(0,0)) + 
  theme_classic() + 
  theme(aspect.ratio = 0.2)
ggsave("data/Figure/Figure2/FigureS3/Rest.pdf", width = 7.5, height = 1.5, units = "in")

varlist <- c("Others", "Nest", "Toilet", "Garbage")
for(var in varlist){
  d <- dta_rel_umap_meansd_l %>% filter(grepl(pattern = var, x = Variable))
  ggplot(d, aes(x = Variable, y = Mean, color = Cluster)) + 
    geom_errorbar(aes(ymax = Mean + SD, ymin = Mean - SD), 
                  position = position_dodge(width=0.8), width = 0.5) + 
    geom_point(position = position_dodge(width=0.8), size = 3) + 
    scale_color_manual(values = colorlist) + 
    scale_x_discrete(expand = c(0,0)) + 
    theme_classic() + 
    theme(aspect.ratio = 0.12)
  ggsave(paste0("data/Figure/Figure2/FigureS3/", var, ".pdf"), width = 12.5, height = 1.5, units = "in")
}
```


##Individual attribute effects

###Data
```{r}
#attribute data
ID_att <- read.csv("data/pre_info_data/individual_information.csv") %>% mutate(Birth = as.Date(Birth))
id <- ID_att$ID

#observation day
obs_day <- read.csv("data/pre_info_data/observation_day.csv") %>% mutate(Obs_day = as.Date(Obs_day))

#add attribute data
dta_rel_umap_att <- dta_rel_umap %>% 
  left_join(ID_att, by = join_by(Colony, ID)) %>% 
  mutate(ID = factor(ID, levels = id)) %>% 
  left_join(obs_day, by = join_by(Colony, Day)) %>% 
  mutate(Birth = as.Date(Birth), 
         Obs_day = as.Date(Obs_day),
         Age_obs = as.integer(Obs_day - Birth),
         log_Age = log(Age_obs),
         Cluster = factor(Cluster), 
         Colony = factor(Colony))
```

###Colony (Figure 2H)
```{r}
#data
dta_rel_umap_att_col <- dta_rel_umap_att %>% 
  group_by(Colony, Cluster) %>% summarise(n = n()) %>% 
  mutate(sum = sum(n), p = n / sum)

#plot
colorlist <- c("#BE5378","#CE7754","#DFC44E","#54B345","#3B98B2","#5F7EB2","#808080")
ggplot(dta_rel_umap_att_col, aes(x=Colony, y=p, fill=Cluster)) + 
  geom_bar(stat = "identity", width = 0.75) +
  scale_y_continuous(limits = c(0,1), expand = c(0,0)) +
  scale_fill_manual(values = colorlist) + 
  theme_classic() + 
  theme(aspect.ratio = 1.5, legend.position = "none")
ggsave("data/Figure/Figure2/Figure2H.pdf", width = 4, height = 6, units = "in")
```

###Caste effect (Figure 2I)
```{r}
#model
ml_caste <- multinom(data = dta_rel_umap_att, Cluster ~ Caste)
glance(ml_caste)
#edf deviance AIC nobs
#12	10030.46	10054.46	3030
Anova(ml_caste, test.statistic = "Chisq")
##x2=1309.5, Df=6, p<0.001

#data
dta_rel_umap_att_caste <- dta_rel_umap_att %>% 
  group_by(Caste, Cluster) %>% summarise(n = n()) %>% 
  mutate(p = n/sum(n)) %>% ungroup()
dta_rel_umap_att_caste <- dta_rel_umap_att_caste %>% 
  add_row(Caste = "B", Cluster = "2", n = 0, p = 0)

#plot
ggplot(dta_rel_umap_att_caste, aes(x=Cluster, y=p, fill=Caste)) + 
  geom_bar(position = position_dodge(), stat = "identity") +
  theme_classic() + 
  scale_y_continuous(expand = c(0,0)) +
  scale_fill_manual(values = c("#8858AA","grey")) + 
  theme(aspect.ratio = 0.5, legend.position = "none")
ggsave("data/Figure/Figure2/Figure2I/bar.pdf", width = 5, height = 2.5, units = "in")

ggplot(dta_rel_umap_att) + 
  geom_point(aes(x=V1, y=V2, color=Caste), size=1) + 
  scale_color_manual(values = c("#8858AA","grey")) + 
  theme_void() + 
  theme(aspect.ratio = 1, legend.position = "none")
ggsave("data/Figure/Figure2/Figure2I/umap.pdf", width = 4, height = 4, units = "in")
```

###Body weight effect (Figure 2J)
```{r}
#model
ml_weight <- multinom(data = dta_rel_umap_att, Cluster ~ Weight)
glance(ml_weight)
#edf deviance AIC nobs
#12	10815.44	10839.44	3030
Anova(ml_weight, test.statistic = "Chisq")
##x2=524.55, Df=6, p<0.001
tidy_ml_weight <- tidy(ml_weight)
write.csv(tidy_ml_weight, row.names = F, "data/statistical_result/Figure2J.csv")

#plot
ggplot(data = dta_rel_umap_att, aes(x = Cluster, color = Cluster, y = Weight)) + 
  geom_violin(position = position_dodge(width = 0.8), width = 0.8) + 
  geom_boxplot(position = position_dodge(0.8), width = 0.125, outlier.shape = NA) + 
  scale_color_manual(values = colorlist) + 
  theme_classic() + 
  theme(aspect.ratio = 0.5, legend.position = "none")
ggsave("data/Figure/Figure2/Figure2J/violin.pdf", width = 5, height = 2.5, units = "in")

spectral <- c("#5E4FA2","#3288BD","#66C2A5","#ABDDA4","#E6F598","#FFFFBF",
              "#FEE08B","#FDAE61","#F46D43","#D53E4F","#9E0142")
ggplot(dta_rel_umap_att) + 
  geom_point(aes(x=V1, y=V2, color=Weight), size=1) + 
  scale_color_gradientn(colors = spectral) + 
  theme_void() + 
  theme(aspect.ratio = 1, legend.position = "none")
ggsave("data/Figure/Figure2/Figure2I/umap.pdf", width = 4, height = 4, units = "in")
```

###Age effect (Figure 2K)
```{r}
#model
ml_age <- multinom(data = dta_rel_umap_att, Cluster ~ log_Age)
glance(ml_age)
#edf deviance AIC nobs
#12	10249.37	10273.37	3030
Anova(ml_age, test.statistic = "Chisq")
##x2=1090.6, Df=6, p<0.001
tidy_ml_age <- tidy(ml_age)
write.csv(tidy_ml_age, row.names = F, "data/statistical_result/Figure2K.csv")

#plot
ggplot(data = dta_rel_umap_att, aes(x = Cluster, color = Cluster, y = log_Age)) + 
  geom_violin(position = position_dodge(width = 0.8), width = 1) + 
  geom_boxplot(position = position_dodge(0.8), width = 0.125, outlier.shape = NA) + 
  scale_y_continuous(breaks = c(log(91), log(182), log(365), log(730), log(1460), log(2920)),
                     labels = c(91, 182, 365, 730, 1460, 2920)) + 
  scale_color_manual(values = colorlist) + 
  theme_classic() + 
  theme(aspect.ratio = 0.5, legend.position = "none")
ggsave("data/Figure/Figure2/Figure2K/violin.pdf", width = 5, height = 2.5, units = "in")

ggplot(dta_rel_umap_att) + 
  geom_point(aes(x=V1, y=V2, color=log_Age), size=1) + 
  scale_color_gradientn(colors = spectral) + 
  theme_void() + 
  theme(aspect.ratio = 1, legend.position = "none")
ggsave("data/Figure/Figure2/Figure2K/umap.pdf", width = 4, height = 4, units = "in")
```

###NB data
```{r}
#NB data
dta_rel_umap_att_nb0 <- dta_rel_umap_att %>% 
  filter(Caste == "N") %>% 
  #relative weight, age
  group_by(Colony) %>% 
  mutate(zAge = scale(log_Age), zWeight = scale(Weight))

#NB & Cluster1-6 data
dta_rel_umap_att_nb_ex7 <- dta_rel_umap_att_nb0 %>% 
  filter(Cluster != "7") %>% 
  mutate(Cluster = factor(Cluster, levels = c("1","2","3","4","5","6")))

#model plot function
eff_pv <- function(eff){
  ggplot(eff) +
    geom_line(aes(x = x, y = predicted, color = response.level)) + 
    scale_color_manual(values = colorlist) + 
    scale_x_continuous(expand = c(0,0)) + 
    scale_y_continuous(expand = c(0,0), limits = c(0,NA)) + 
    theme_classic() + 
    theme(aspect.ratio = 0.5, legend.position = "none")
}
```

####Relative body weight effect (Figure 2L)
```{r}
#model
ml_nb_zwei2 <- multinom(
  data = dta_rel_umap_att_nb_ex7, Cluster ~ poly(zWeight,2)
)
glance(ml_nb_zwei2)
#edf deviance AIC nobs
#15	8858.11	8888.11	2686	
Anova(ml_nb_zwei2, test.statistic = "Chisq")
##x2=394.22, Df=10, p<0.001
tidy_ml_nb_zwei2 <- tidy(ml_nb_zwei2)
write.csv(tidy_ml_nb_zwei2, row.names = F, "data/statistical_result/Figure2L.csv")

#plot
eff <- ggeffect(ml_nb_zwei2, terms = "zWeight[all]")
eff_pv(eff)
ggsave("data/Figure/Figure2/Figure2L.pdf", width = 4, height = 2, units = "in")
```

####Relative age (Figure 2M)
```{r}
#model
ml_nb_zage2 <- multinom(
  data = dta_rel_umap_att_nb_ex7, Cluster ~ poly(zAge,2)
)
glance(ml_nb_zage2)
#edf deviance AIC nobs
#15	8978.555	9008.555	26866	
Anova(ml_nb_zage2, test.statistic = "Chisq")
##x2=273.78, Df=10, p<0.001
tidy_ml_nb_zage2 <- tidy(ml_nb_zage2)
write.csv(tidy_ml_nb_zage2, row.names = F, "data/statistical_result/Figure2M.csv")

#plot
eff <- ggeffect(ml_nb_zage2, terms = "zAge[all]")
eff_pv(eff)
ggsave("data/Figure/Figure2/Figure2M.pdf", width = 4, height = 2, units = "in")
```

####Sex (Figure S2B)
```{r}
#model
ml_nbsex <- multinom(data = dta_rel_umap_att_nb_ex7, Cluster ~ Sex)
glance(ml_nbsex)
#edf deviance AIC nobs
#10	9221.68	9241.68	2686
Anova(ml_nbsex, test.statistic = "Chisq")
##x2=30.653, Df=5, p<0.001

#data
dta_rel_umap_att_castesex <- dta_rel_umap_att %>% 
  group_by(Sex, Cluster) %>% summarise(n = n()) %>% ungroup() %>% 
  pivot_wider(names_from = Cluster, values_from = n) %>% 
  pivot_longer(cols = -1, names_to = "Cluster", values_to = "n") %>% 
  replace_na(list(n = 0)) %>% arrange(Sex, Cluster) %>% 
  group_by(Sex) %>% mutate(p = n / sum(n)) %>% ungroup()

#plot
ggplot(dta_rel_umap_att_castesex, aes(x=Cluster, y=p, fill=Sex)) + 
  geom_bar(position = position_dodge(), stat = "identity") +
  theme_classic() + 
  scale_y_continuous(expand = c(0,0)) +
  scale_fill_manual(values = c("#D04255","#6CBAD8","#EDA184","#1DAF9E")) + 
  theme(aspect.ratio = 0.5, legend.position = "none")
ggsave("data/Figure/Figure2/FigureS2B/bar.pdf", width = 5, height = 2.5, units = "in")

ggplot(dta_rel_umap_att) + 
  geom_point(aes(x=V1, y=V2, color=Sex), size=1) + 
  scale_color_manual(values = c("#D04255","#6CBAD8","#EDA184","#1DAF9E")) + 
  theme_void() + 
  theme(aspect.ratio = 1, legend.position = "none")
ggsave("data/Figure/Figure2/FigureS2B/umap.pdf", width = 4, height = 4, units = "in")
```

####Sex * relative body weight (Figure S2C)
```{r}
#model
ml_nb_sex_zwei2 <- multinom(
  data = dta_rel_umap_att_nb_ex7, Cluster ~ Sex * poly(zWeight,2)
)
glance(ml_nb_sex_zwei2)
#edf deviance AIC nobs
#30	8738.96	8798.96	2686	
Anova(ml_nb_sex_zwei2, test.statistic = "Chisq")
##x2=102.87, Df=10, p<0.001
#LR Chisq Df Pr(>Chisq) 
#Sex                     16.28  5   0.006081 ** 
#poly(zWeight, 2)       379.85 10  < 2.2e-16 ***
#Sex:poly(zWeight, 2)   102.87 10  < 2.2e-16 ***
tidy_ml_nb_sex_zwei2 <- tidy(ml_nb_sex_zwei2)
write.csv(tidy_ml_nb_sex_zwei2, row.names = F, "data/statistical_result/FigureS2C.csv")

#plot
ml_nbf_zwei2 <- multinom(
  data = dta_rel_umap_att_nb_ex7 %>% filter(Sex == "F"), Cluster ~ poly(zWeight,2)
)
eff <- ggeffect(ml_nbf_zwei2, terms = "zWeight[all]")
eff_pv(eff)
ggsave("data/Figure/Figure2/FigureS2C/NF.pdf", width = 4, height = 3, units = "in")

ml_nbm_zwei2 <- multinom(
  data = dta_rel_umap_att_nb_ex7 %>% filter(Sex == "M"), Cluster ~ poly(zWeight,2)
)
eff <- ggeffect(ml_nbm_zwei2, terms = "zWeight[all]")
eff_pv(eff)
ggsave("data/Figure/Figure2/FigureS2C/NM.pdf", width = 4, height = 3, units = "in")
```

####Sex * relative age (Figure S2D)
```{r}
#model
ml_nb_sex_zage2 <- multinom(
  data = dta_rel_umap_att_nb_ex7, Cluster ~ Sex * poly(zAge,2)
)
glance(ml_nb_sex_zage2)
#edf deviance AIC nobs
#30	8800.866	8860.866	2686	
Anova(ml_nb_sex_zage2, test.statistic = "Chisq")
##x2=155.42, Df=10, p<0.001
#LR Chisq Df Pr(>Chisq)
#Sex                 22.273  5  0.0004646 ***
#poly(zAge, 2)      265.397 10  < 2.2e-16 ***
#Sex:poly(zAge, 2)  155.416 10  < 2.2e-16 ***
tidy_ml_nb_sex_zage2 <- tidy(ml_nb_sex_zage2)
write.csv(tidy_ml_nb_sex_zage2, row.names = F, "data/statistical_result/FigureS2D.csv")

#plot
ml_nbf_zage2 <- multinom(
  data = dta_rel_umap_att_nb_ex7 %>% filter(Sex == "F"), Cluster ~ poly(zAge,2)
)
eff <- ggeffect(ml_nbf_zage2, terms = "zAge[all]")
eff_pv(eff)
ggsave("data/Figure/Figure2/FigureS2D/NF.pdf", width = 4, height = 3, units = "in")

#zAge2 Mのみ
ml_nbm_zage2 <- multinom(
  data = dta_rel_umap_att_nb_ex7 %>% filter(Sex == "M"), Cluster ~ poly(zAge,2)
)
eff <- ggeffect(ml_nbm_zage2, terms = "zAge[all]")
eff_pv(eff)
ggsave("data/Figure/Figure2/FigureS2D/NM.pdf", width = 4, height = 3, units = "in")
```


#Figure 3

##Cluster assingnment (Figures 3A & 3B)
```{r}
#each individual & each day
day <- dayname %>% unlist()
dta_rel_umap_day <- dta_rel_umap %>% 
  mutate(Day = factor(Day, levels = day)) %>% 
  arrange(Colony, Day) %>% 
  group_by(Colony, ID) %>% 
  mutate(Day = factor(as.character(1:30), levels = rev(as.character(1:30))))

sizelist <- c(22, 17, 19, 21, 22)
colorlist <- c("#BE5378","#CE7754","#DFC44E","#54B345","#3B98B2","#5F7EB2","#808080")

for(i in 1:5){
  colony <- colonyname[i]; size <- sizelist[i]
  ggplot(dta_rel_umap_day %>% filter(Colony == colony)) + 
    geom_tile(aes(x=ID, y=Day, fill=Cluster)) +
    scale_fill_manual(values = colorlist) +
    scale_y_discrete(expand = c(0,0)) +
    scale_x_discrete(expand = c(0,0)) +
    theme_classic() + 
    theme(aspect.ratio = 22/size, legend.position = "none")
  ggsave(paste0("data/Figure/Figure3/Figure3A/", colony, ".pdf"), 
         width = size*0.3, height = 6.6, units = "in")
}

#each individual & across 30 days
cluster <- c("1","2","3","4","5","6","7")
dta_rel_umap_id <- dta_rel_umap %>% 
  group_by(Colony, ID, Cluster) %>% 
  summarise(n = n()) %>% 
  pivot_wider(names_from = Cluster, values_from = n) %>% 
  pivot_longer(cols = c(-1,-2), names_to = "Cluster", values_to = "n") %>% 
  replace_na(list(n = 0)) %>% 
  mutate(ID = factor(ID, levels = idvect),
         Cluster = factor(Cluster, levels = cluster)) %>% 
  arrange(Colony, ID, Cluster)

for(i in 1:5){
  colony <- colonyname[i]; size <- sizelist[i]
  for(j in 1:7){
    clu <- cluster[j]; col <- colorlist[j]
    d <- dta_rel_umap_id %>% filter(Colony == colony, Cluster == clu)
    ggplot(d) + 
      geom_tile(aes(x=ID, y=Cluster, fill=n)) +
      theme_void() + 
      scale_fill_gradient(limits = c(0, 30), low = "white", high = col) +
      scale_y_discrete(expand = c(0,0)) +
      scale_x_discrete(expand = c(0,0)) +
      theme(aspect.ratio = 1.54/size, legend.position = "none")
    ggsave(paste0("data/Figure/Figure3/Figure3B/", colony, "_", clu, ".pdf"), 
           width = size*0.3, height = 0.55, units = "in")
  }
}
```

##Representative cluster for each individual (for Figures 4-6)
```{r}
dta_rel_umap_clu <- dta_rel_umap %>% 
  group_by(Cluster) %>% summarise(n = n()) %>% mutate(p = n / sum(n)) %>% select(-n)

dta_rel_umap_id_max <- dta_rel_umap_id %>% 
  mutate(Cluster = as.integer(Cluster)) %>% 
  left_join(dta_rel_umap_clu, join_by(Cluster)) %>% 
  group_by(Colony, ID) %>% 
  filter(n >= max(n), p <= min(p)) %>% 
  mutate(p = n/30) %>% ungroup() %>% 
  mutate(Cluster = as.character(Cluster))

write.csv(dta_rel_umap_id_max, row.names = F, 
          "data/processed_data/representative_cluster.csv")
```


##Consistency index (Figure 3C)
```{r}
#cluster assignment across all individuals
dta_rel_umap_id_cluster_mean_all <- dta_rel_umap_id %>% 
  ungroup() %>% group_by(Cluster) %>% summarise(p = sum(n) / (n() * 30))

#consistency index for observed data
dta_rel_umap_id_cons <- dta_rel_umap_id %>% 
  left_join(dta_rel_umap_id_cluster_mean_all) %>% 
  mutate(logp = if_else(n == 0, 1e-10, log(n/30)),
         ent = n/30 * logp) %>% 
  summarise(Consistency = 1 + sum(ent) / log(7))

#consistency index for randomized data

##overall assignment rate for each cluster across individuals
clu <- dta_rel_umap_id_cluster_mean_all %>% mutate(cp = cumsum(p)) %>% pull(cp)

##randomized data (10000 sets)
set.seed(10)
ran <- runif(300000, min=0, max=1)
randata <- tibble(ran) %>% 
  mutate(No = rep(1:10000, each = 30)) %>% 
  mutate(Cluster = case_when(
    ran <= clu[1] ~ "1",
    ran > clu[1] & ran <= clu[2] ~ "2",
    ran > clu[2] & ran <= clu[3] ~ "3", 
    ran > clu[3] & ran <= clu[4] ~ "4", 
    ran > clu[4] & ran <= clu[5] ~ "5", 
    ran > clu[5] & ran <= clu[6] ~ "6", 
    ran > clu[6] ~ "7"))
randata_cons <- randata %>% 
  group_by(No, Cluster) %>% 
  summarise(n = n()) %>% 
  left_join(dta_rel_umap_id_cluster_mean_all) %>% 
  mutate(logp = if_else(n == 0, 1e-10, log(n/30)), 
         ent = n/30 * logp) %>% 
  summarise(Consistency = 1 + sum(ent) / log(7))
  mutate(Cons_rank = 10001 - row_number(Consistency)) %>% 
  arrange(Cons_rank)

#calculate p-value (right)
dta_rel_umap_id_cons$p_value <- sapply(dta_rel_umap_id_cons$Consistency, function(x) {
  mean(randata_cons$Consistency >= x)
})
dta_rel_umap_id_cons <- dta_rel_umap_id_cons %>% 
  mutate(Sig = case_when(
    p_value < 0.001 ~ "***", p_value >= 0.001 & p_value < 0.01 ~ "**",
    p_value >= 0.01 & p_value < 0.05 ~ "*", p_value >= 0.05 ~ "NS"
    ))
dta_rel_umap_id_cons
write.csv(dta_rel_umap_id_cons, row.names = F, "data/statistical_result/Figure3C.csv")

#plot
val95 <- randata_cons %>% filter(Cons_rank == 500) %>% pull(Consistency)
for(colony in colonyname){
  d <- dta_rel_umap_id_cons %>% filter(Colony == colony)
  size <- d %>% nrow()
  ggplot() + 
    geom_bar(data=d, aes(x = ID, y = Consistency^0.5), stat = "identity") +
    geom_hline(yintercept = val95^0.5, linetype = "dashed") + 
    theme_classic() + 
    scale_y_continuous(limits = c(0,1), expand = c(0,0)) +
    theme(aspect.ratio = 5.5/size, axis.title = element_blank(), axis.text = element_blank())
  ggsave(paste0("data/Figure/Figure3/Figure3C/", colony, ".pdf"), 
         width = size*0.2, height = 1.25, units = "in")
}
```


##Cosine similarity (Figure 3D)
```{r}
#cosine similarity function
cossim <- function(a,b){sum(a*b) / (sqrt(sum(a)) * sqrt(sum(b)))}
dta_rel_umap
#calculate cosine similarity in observed data
d_cossim <- dta_rel_umap %>% 
  mutate(n = 1) %>% 
  pivot_wider(names_from = Cluster, values_from = n) %>% 
  select(Colony, ID, Day, "1","2","3","4","5","6","7") %>% 
  pivot_longer(cols = 4:10, names_to = "Cluster", values_to = "n") %>% 
  replace_na(list(n = 0)) %>% 
  group_by(Colony, Cluster, Day) %>% 
  nest() %>% 
  arrange(Colony, Cluster, Day) %>% 
  ungroup() %>% group_by(Colony, Cluster) %>% 
  mutate(data_lead = lead(data)) %>% 
  unnest() %>% 
  select(-ID1) %>% ungroup() %>% 
  group_by(Colony, Cluster, Day) %>% nest() %>% 
  mutate(cos_sim = map_dbl(data, ~ cossim(.$n, .$n1))) %>% 
  ungroup() %>% 
  select(Colony, Cluster, cos_sim) %>% 
  mutate(across(cos_sim, ~ replace(., is.nan(.), 0))) %>% 
  drop_na(everything())

#permuted data
##randomly permute individual clusters per colony & day
##permuted vectors (30 days * 100 sets * 5 colonies)
##calculate cosine similarities (29 * 100 sets * 5 colonies * 7 clusters)
for(colony in colonyname){
  d <- dta_rel_umap %>% filter(Colony == colony)
  Cluster <- d %>% pull(Cluster)
  N <- length(Cluster)
  
  set.seed(20)
  for(i in 1:100){
    perm <- shuffle(N)
    Cluster_perm <- Cluster[perm]
    d_perm <- d %>% mutate(Cluster = Cluster_perm)
    
    d_perm_res <- d_perm %>% 
      mutate(n = 1) %>% 
      pivot_wider(names_from = Cluster, values_from = n) %>% 
      pivot_longer(cols = 4:10, names_to = "Cluster", values_to = "n") %>% 
      replace_na(list(n = 0)) %>% 
      group_by(Colony, Cluster, Day) %>% 
      nest() %>% 
      arrange(Colony, Cluster, Day) %>% 
      ungroup() %>% group_by(Colony, Cluster) %>% 
      mutate(data_lead = lead(data)) %>% 
      unnest() %>% 
      select(-ID1) %>% ungroup() %>% 
      group_by(Colony, Cluster, Day) %>% nest() %>% 
      mutate(cos_sim = map_dbl(data, ~ cossim(.$n, .$n1))) %>% 
      mutate(across(cos_sim, ~ replace(., is.nan(.), 0))) %>% 
      ungroup %>% select(Colony, Cluster, cos_sim)
    
    d_perm0 <- d_perm0 %>% bind_rows(d_perm_res)
  }
}
d_perm <- d_perm0 %>% slice(-1) %>% arrange(Colony, Cluster) %>% drop_na(everything())

#summarize permuted data
d_cossim_perm_m <- d_perm %>% 
  ungroup() %>% group_by(Colony, Cluster) %>% 
  summarise(med_perm = median(cos_sim), mean_perm = mean(cos_sim), 
            sd_perm = sd(cos_sim), se_perm = sd(cos_sim)/sqrt(n()))

#integrate observed data and permuted data
d_perm <- d_perm %>% mutate(actperm = "Perm")
d_cossim_actperm <- d_cossim %>% mutate(actperm = "Act") %>% bind_rows(d_perm)
d_cossim_actperm_m <- d_cossim_m %>% 
  left_join(d_cossim_perm_m, by = join_by(Colony, Cluster))

#plot
ggplot(d_cossim_actperm_m) + 
  geom_errorbar(aes(x=Cluster, y=mean, group=Colony, ymax = mean + se, ymin = mean - se*0.2), 
                position = position_dodge(width=0.75), width = 0.4) +
  geom_bar(aes(x=Cluster, y=mean, fill=Colony), 
           position = position_dodge(width=0.75), stat = "identity", width = 0.75) + 
  geom_point(aes(x=Cluster, y=mean_perm, color=Colony), 
             position = position_dodge(width=0.75)) + 
  scale_y_continuous(expand = c(0,0)) + 
  scale_fill_grey() + 
  scale_color_manual(values = rep("white",5)) + 
  theme_classic() + 
  theme(aspect.ratio = 0.32, legend.position = "none", 
        axis.title = element_blank(), axis.text = element_blank())
ggsave("data/Figure/Figure3/Figure3D.pdf", width = 6, height = 2, units = "in")

#test
colony <- "0"; cluster <- "0";
p_value <- 0
d0 <- tibble(colony, cluster, p_value)
for(colony in colonyname){
  for(cluster in clustername){
    actual_vector <- d_cossim %>% filter(Colony == colony, Cluster == cluster) %>% pull(cos_sim)
    perm_vector <- d_perm %>% filter(Colony == colony, Cluster == cluster) %>% pull(cos_sim)
    wt <- wilcox.test(actual_vector, perm_vector)
    f_value <- wt$statistic
    p_value <- wt$p.value
    d <- tibble(colony, cluster, f_value, p_value)
    d0 <- d0 %>% bind_rows(d)
  }
}
d_cossim_wtest <- d0 %>% slice(-1) %>% 
  mutate(sig = case_when(
    p_value >= 0.05 ~ "NS", p_value < 0.05 & p_value >= 0.01 ~ "*", 
    p_value < 0.01 & p_value >= 0.001 ~ "**", p_value < 0.001 ~ "***" 
  )) %>% 
  arrange(cluster)
write.csv(d_cossim_wtest, row.names = F, "data/statistical_result/Figure3D.csv")
```


##Co-assignment index (Figure 3E)
```{r}
#data
dta_rel_umap_id_1 <- dta_rel_umap_id %>% mutate(p = n/30) %>% select(Colony, ID, Cluster1 = Cluster, p1 = p)
dta_rel_umap_id_2 <- dta_rel_umap_id %>% mutate(p = n/30) %>% select(Colony, ID, Cluster2 = Cluster, p2 = p)

dta_rel_umap_id_cluster_mean_all_1 <- dta_rel_umap_id_cluster_mean_all %>% select(Cluster1 = Cluster, f1 = p)
dta_rel_umap_id_cluster_mean_all_2 <- dta_rel_umap_id_cluster_mean_all %>% select(Cluster2 = Cluster, f2 = p)

dta_rel_umap_id_12 <- dta_rel_umap_id_1 %>% 
  ##empirical co-assignment strength
  left_join(dta_rel_umap_id_2, by = join_by(Colony, ID)) %>% 
  mutate(cooc_obs = if_else(Cluster1 == Cluster2, p1*p2, p1*p2*2)) %>% 
  ##reference co-assignment strength
  left_join(dta_rel_umap_id_cluster_mean_all_1, by = join_by(Cluster1)) %>% 
  left_join(dta_rel_umap_id_cluster_mean_all_2, by = join_by(Cluster2)) %>% 
  mutate(cooc_exp = if_else(Cluster1 == Cluster2, f1*f2, f1*f2*2)) %>% 
  ##odds ratio
  mutate(cooc_ind = cooc_obs / cooc_exp) %>% 
  unite(Comb, Cluster1, Cluster2, remove = F, sep = "_")

#summarized data
dta_rel_umap_id_12_s <- dta_rel_umap_id_12 %>% 
  group_by(Cluster1, Cluster2, Comb) %>% 
  summarise(mean_cooc_ind = mean(cooc_ind), sd_cooc_ind = sd(cooc_ind))

#plot
ggplot(dta_rel_umap_id_12_s, aes(x = Cluster1, y = sqrt(mean_cooc_ind), fill = Cluster2)) + 
  geom_bar(stat = "identity", position = position_dodge(width = 0.9)) + 
  scale_y_continuous(expand = c(0,0)) + 
  scale_fill_manual(values = colorlist) + 
  theme_classic() + 
  theme(aspect.ratio = 0.25, legend.position = "none")
ggsave("data/Figure/Figure3/Figure3E.pdf", width = 10, height = 2.5, units = "in")
```


